### 2022-05-06 Report/paper.tex
initial import of the tex source

\documentclass{article}

\usepackage{graphicx}
\usepackage{pdflscape}
\usepackage{afterpage}
\extrafloats{200}
\maxdeadcycles=1000
% if you need to pass options to natbib, use, e.g.:
%     \PassOptionsToPackage{numbers, compress}{natbib}
% before loading neurips_2021

% ready for submission
\usepackage[final]{neurips_2021}

% to compile a preprint version, e.g., for submission to arXiv, add add the
% [preprint] option:
%     \usepackage[preprint]{neurips_2021}

% to compile a camera-ready version, add the [final] option, e.g.:
%     \usepackage[final]{neurips_2021}

% to avoid loading the natbib package, add option nonatbib:
%    \usepackage[nonatbib]{neurips_2021}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{xcolor}         % colors

\title{Results from the deep learning challenge associated with the 3MD4040 lecture on plankton classification}

% The \author macro works with any number of authors. There are two commands
% used to separate the names and addresses of multiple authors: \And and \AND.
%
% Using \And between authors leaves it to LaTeX to determine where to break the
% lines. Using \AND forces a line break at that point. So, if LaTeX puts 3 of 4
% authors names on the first line, and the last on the second line, try using
% \AND instead of \And before the third author name.

\author{%
Jérémy Fix\\
\textit{LORIA, CNRS,} \\
\textit{Université de Lorraine, CentraleSupélec}\\
Gif-sur-Yvette, France \\
\texttt{jeremy.fix@centralesupelec.fr}
}

\begin{document}

\maketitle

\begin{abstract}
	This year, the deep learning lecture on the Metz campus of CentraleSupélec was evaluated by the participation to a custom challenge on plankton classification hosted on Kaggle. This
	paper presents the competition and the results obtained by the 13 "competing"
	teams. The challenge was hosted on kaggle \url{https://www.kaggle.com/c/3md4040-2022-challenge}.
\end{abstract}

\section{Context}

\subsection{Why a challenge}

I used to evaluate the deep learning lecture by a standard paper and pen exam. But I always thought that was not the best way to evaluate the students because deep learning is a know-how in addition to a theoretical understanding of the concepts. You cannot obtain reasonnable performance with your deep learning code if you do not understand what is going on. There are so many degrees of freedom that a random exploration of the hyperparameters is hopeless.

The challenge lasted approximately 2 months, started in December 2021 and ended beginning of February 2022.

\subsection{Grading}

Grading of the students' contributions is on three topics :
\begin{itemize}
	\item the performance on the leaderboard for 5 points
	\item the quality of the 20 minutes long video recorded presentation for 5 points
	\item the quality of the code and experimentations for 10 points
\end{itemize}

I once asked orally to provide a little article presenting the experimental track but forgot to mention it on the grading page and therefore did not ask for it at the end of the challenge.

For the performance on the leaderboard, the points were allocated as~:
\begin{itemize}
	\item 1 point if the submitted predictor performs better than a benchmark linear network
	\item 2 points if the submitted predictor performas better than a benchmark basic CNN network
	\item 2 points if the submitted predictor ranks first
\end{itemize}

\subsection{Rules}

The students were asked to respect the following rules :
\begin{itemize}
	\item the predictor must be a neural network
	\item the code must be in pytorch. Using higher level framework such as pytorch lightning was allowed
	\item the students were not allowed to share codes between the teams
	\item the students were allowed and strongly encouraged to ask questions on a dedicated forum and they would be provided hints under the condition that the previous rules were not violated
\end{itemize}

In practice, video sessions were organized with the groups that had the most difficulties to come into the challenge.

%-----------------------------------------------------------------------------------
\section{The challenge on plankton classification}

The data for the challenge come from the ZooScanNet dataset (\url{https://www.seanoe.org/data/00446/55741/}). It contains 1,433,278 images sorted in 93 taxa. We only used the images and not the native nor skimage features that were provided. It does not mean that these handcrafted features could not be useful for plankton classification, it is just the the challenge focused exclusively on image classifcation (but definitely, a later issue should include them).

\subsection{Data preparation}

Three folds were prepared from the original data :
\begin{itemize}
	\item a training set of 855.394 images
	\item a public test set of 47.473 images
	\item a private test set of 47.472 images
\end{itemize}

We did not kept all the classes from the original 93 taxa. Five classes were ignored : \emph{badfocus\_artefact}, \emph{badfocus\_Copepoda}, \emph{bubble}, \emph{multiple\_\_Copepoda} and \emph{multiple\_\_other}. Three classes were fused : \emph{detritus}, \emph{fiber\_\_detritus} and \emph{seaweed}. Therefore, at the end, we get a total of $86$ classes. We also kept a maximum of $200.000$ per classes (before merging classes which explains why the detritus has more than $200000$ samples).

One concern in preparing the data challenge is to prevent cheating such as overfitting the original dataset (which then comprise the test data). To prevent the challengers from cheating, a random rotation was applied to the image.

\subsection{Data statistics}

All the sets were identically balanced, i.e. the classes are imbalanced but imbalanced in the same proportions irrespectively of the fold. After the challenge, rereading the different classes, it might have been worth fusing artefact with detritus, and also maybe some egg classes. There are also two classes for tail and head which might gather tails and heads of different species which in no way can be considered as a single class. These two classes should have probably been discarded.

\begin{figure}
	\includegraphics[width=\columnwidth]{figs/stats.pdf}
	\caption{\label{fig:classdistrib} The distribution of samples per classes. See the appendix for the details, especially the class names}
\end{figure}

The distribution of the samples over the different folds is shown on figure~\ref{fig:classdistrib}. The dataset is imbalanced and that was one interest of the challenge. The class with the most samples is the detritus followed by the Calanoida. The class with the lowest number of samples is the Ctenophora which has only $36$ samples in the training set and $1$ sample in both the public and private test sets.

The images vary in size. During the prepation of the dataset, the images were resized so that their height and width do not exceed $300$, keeping their aspect ratio. Athough the original images vary in sizes, this choice was made to keep the weight of the whole dataset reasonnable. Note however that this choice may have introduced difficulties in the recognition process (which could have been mitigated by providing the handcrafted features). Indeed, this resize erases class specific shape distributions. The distribution of heights and widhts of the datasets are shown on figure~\ref{fig:sizes}, left for the training set, \ref{fig:sizes}, middle for the test set and \ref{fig:sizes}, right for the original images. For practical simplicity, only the size plot of the training set indicates the class belongings.

\begin{figure}
	\begin{center}
		% \includegraphics[width=0.3\columnwidth]{figs/size_train.png}
		% \includegraphics[width=0.3\columnwidth]{figs/sizes_test.pdf}
		% \includegraphics[width=0.3\columnwidth]{figs/size_orig.png}
	\end{center}
	\caption{\label{fig:sizes} Distribution of the image sizes for the training set (left), test set(right) and the original dataset (right). On the plot on the right, the original sizes have been limited to $4000$ for both widths and heights but there exists outliers up to $10000$ in width or $12000$ in height.}
\end{figure}

%TODO: Plot size distribution as a 2D scatter plot

\subsection{Evaluation metrics}

The challenge is a classification challenge with strongly imbalanced classes. The competitors were evaluated with the macro-average F1 score which is the average of the class F1 scores.

The class F1 score is defined as the harmonic mean of the precision and recall. Given a class $k$, the number of true positives $TP_k$, false positive $FP_k$, and false negatives $FN_k$, precision, recall and $F1$ are computed as~:
\begin{eqnarray*}
	precision_k &=& \frac{TP_k}{TP_k + FP_k}\\
	recall_k &=& \frac{TP_j}{TP_k + FN_k}\\
	\frac{1}{F1_k} &=& \frac{1}{2}(\frac{1}{precision_k} + \frac{1}{recall_k})
\end{eqnarray*}

The macro-average F1 is then defined as the average of all the classes~:

\begin{equation}
	\mbox{macro-F1} = \frac{1}{86}\sum_{k=0}^{85} F1_k
\end{equation}

%-----------------------------------------------------------------------------------
\section{The submissions of the participants}

In the next section, we give a quick overview of the different techniques involved by the participants for the different topics of :

\begin{itemize}
	\item data loading 
	\item data augmentation
	\item classification model architectures
	\item optimization setup (optimizer, scheduler, ..)
	\item handling of the class imbalance
\end{itemize}

That overview is built from inspection of the submitted codes and may not reflect all the configurations that have been tested by the participants. The collected information from the code are summarized in tables~\ref{table:overview1},\ref{table:overview2},\ref{table:overview3}.

\afterpage{
	\clearpage
	\thispagestyle{empty}
	\begin{landscape}
		\begin{table}
			\begin{tabular}{p{3cm}|p{3cm}|p{3cm}|p{3cm}|p{3cm}|p{3cm}|p{3cm}}
			Participant & Data preprocessing & Train augmentation & Test augmentation & Model architectures & Optimizer and scheduler & Class imbalance \\
			\hline
			SpongeBob & 
			Resize($224$), Normalize($0.5$, $0.5$) & 
			HFlip($0.5$), VFlip($0.5$) , Rotate($-40, 40$), Translate($0.3, 0.5)$, Scale($0.7, 1.3$), Shear($-30, 30$)&
			None &
			Custom ConvNet, Pretrained torchvision ResNet, AlexNet, DenseNet121, timm ViT, EfficientNetB2, ConvNeXt, RegNetx\_320, EfficientNetv2\_rw\_t, BeIT, CoaT&
			Cross Entropy loss, Adam(lr=$5e-4$), ReduceLROnPlateau(validation loss), early stopping on validation F1, Batch size $64$,  Random train/valid split ($0.95, 0.05$)&
			Class weights in the CE loss\\
			\hline
			DeepWhale & 
			Pad/Resize($224$), ImageNet normalization&
			HFlip($0.5$), VFlip($0.5$), Rotate($-150, 150$)&
			None& 
			Custom CNN, pretrained ResNet50, 101, 152, VGG19, SqueezeNet from torchvision&
			Cross EntropyLoss, SGD($1e-3$), ReduceLROnPlateau(validation loss), early stopping on the validation loss, batch size $32$, train/valid split ($0.8, 0.2$) per class &
			batch sampler with $f_k = 1/count_k$\\
			\hline
			GrandeRegazzoni & 
			SquarePad($300$) or ResizeCrop($128$) or Resize($150$)&
			HFlip($0.5$), VFlip($0.5$), Rotate($-40, 40$), Translate($-0.2, 0.2$)&
			None&
			Custom ConvNet, EfficientNet\footnote{\url{https://github.com/lukemelas/EfficientNet-PyTorch}}, HRNet, ResNet, SwinTransformer&
			Cross Entropy loss, Adam($1e-3$), ReduceLROnPlateau(validation F1), early stopping on the validation loss, Batch size $128$, train/valid split ($0.8, 0.2$) per class &
			batch sampler with $f_k = 1/count_k$\\
			\hline
			Voleurs d'huile&
			Resize($224$), z-score normalization by training mean/std&
			HFlip($0.5$), VFlip($0.5$), Gaussian Blur&
			None&
			Custom CNN, pretrained torchvision ResNet and DenseNet121 & 
			Cross Entropy loss, Adam($4e-4$), StepLR($\gamma=0.1, step=6$), early stopping on the validation loss, Batch size $32$, train/valid split ($0.8, 0.2$) &
			None\\
			\hline
		\end{tabular}
			\caption{Overview of the experiments of the participants. Part 1. \label{table:overview1}}
	\end{table}

		\begin{table}
			\begin{tabular}{p{3cm}|p{3cm}|p{3cm}|p{3cm}|p{3cm}|p{3cm}|p{3cm}}
			Participant & Data preprocessing & Train augmentation & Test augmentation & Model architectures & Optimizer and scheduler & Class imbalance \\
			\hline
			Metz Dream Deep & 
			Resize($224$), Normalize($0.5$, $0.5$) & 
				Rotate($0, 360$), HFlip($0.5$), Translate($0.2$), Normalization$(0.5, 0.5$)&
			None&
			&
				Cross entropy loss, Adam($1e-4$), Weight decay $1e-4$, ..., batch size $128$, train/valid split&
			batch sampler $f_k=1/count_k$\\
			\hline
			The Backpropagated StuBBorns & 
			Resize($224$), Normalize($0.5$, $0.5$) & 
			&
			&
			&
			&
			\\
			\hline
			JMBmc & 
			Resize($224$), Normalize($0.5$, $0.5$) & 
			&
			&
			&
			&
			\\
			\hline
			AG0D & 
			Resize($224$), Normalize($0.5$, $0.5$) & 
			&
			&
			&
			&
			\\
			\hline
		\end{tabular}
			\caption{Overview of the experiments of the participants. Part 2. \label{table:overview2}}
	\end{table}

		\begin{table}
			\begin{tabular}{p{3cm}|p{3cm}|p{3cm}|p{3cm}|p{3cm}|p{3cm}|p{3cm}}
			Participant & Data preprocessing & Train augmentation & Test augmentation & Model architectures & Optimizer and scheduler & Class imbalance \\
			\hline
			ResMetz & 
			Resize($224$), Normalize($0.5$, $0.5$) & 
			&
			&
			&
			&
			\\
			\hline
			CelestineFeuillat & 
			Resize($224$), Normalize($0.5$, $0.5$) & 
			&
			&
			&
			&
			\\
			\hline
			LesJongleurs & 
			Resize($224$), Normalize($0.5$, $0.5$) & 
			&
			&
			&
			&
			\\
			\hline
			DeepBeru4s & 
			Resize($224$), Normalize($0.5$, $0.5$) & 
			&
			&
			&
			&
			\\
			\hline
			The Whale Sharks & 
			Resize($224$), Normalize($0.5$, $0.5$) & 
			&
			&
			&
			&
			\\
		\end{tabular}
			\caption{Overview of the experiments of the participants. Part 3. \label{table:overview3}}
	\end{table}
	\end{landscape}
	\clearpage
}

Some of the participants also considered model averaging :

SpongeBox , GrandeRegazzoni (convenet, hrnet, efficientnet)

%TODO :Verifier le cout du padding : Fill value ?? 0.0 par défaut mais pour du blanc c'est 1.0 ?!

\subsection{The scores of the participants through time}

I find it interesting to see how the submissions of the participants improved through time. The teams submitted from 5 to 79 entries and the scores they obtained on the public and private test sets are shown on figure~\ref{fig:F1scores}.

\begin{figure}
	\begin{center}
		\includegraphics[width=0.45\columnwidth]{figs/public-test.pdf}
		\includegraphics[width=0.45\columnwidth]{figs/private-test.pdf}
	\end{center}
	\caption{\label{fig:F1scores} Macro F1 on the public (left) and private (right) test sets as a function of time.}
\end{figure}

\subsection{Analysis of the results}

The best entries of all the teams ranged from 0.71 to 0.82 (except one outlier at 0.57) on the private leaderboard in macro average F1 score. The class F1 scores, computed from the whole test data (public and private) is displayed on figure~\ref{fig:class_f1_scores}, left. This graph displays the distribution of class F1 scores for each team, where the teams are ordered by decreasing macro average F1.

\begin{figure}
	\begin{center}
		\raisebox{-0.5\height}{\includegraphics[width=0.25\columnwidth]{figs/class_f1_scores_per_team.pdf}}
		\hspace*{.4in}
		\raisebox{-0.5\height}{\includegraphics[width=0.62\columnwidth]{figs/class_f1_scores_per_class.pdf}}
	\end{center}
	\caption{\label{fig:class_f1_scores} Class F1 scores on the test data per team (left) and per class (right).}
\end{figure}

We can also plot the F1 scores from the class perspective rather than a team perspective. That view of the class F1 scores is shown on figure~\ref{fig:class_f1_scores}, right. This graph plots the distribution of the teams' F1 for every class. From this graph, we see that some classes are consistently well predicted (e.g. the classes $0$, $1$, $2$, $3$, $4$ and $6$) and some classes seem much harder to classify ($80$, $76$, $75$, $70$, $66$). In order to get an idea of the classes that appear to be the easiest to predict, we selected from the distribution of F1 scores, the classes for which the minimal F1 over all the teams is higher than 0.8 This choice leads to the easiest classes given in table~\ref{table:best_predicted}.

\input{table_best_predicted.tex}

In order to get an idea of the classes that appear to be the most difficult to predict, we selected from the distribution of F1 scores, the classes for which the maximal F1 over all the teams is smaller than 0.6. This choice leads to the hardest classes given in table~\ref{table:worst_predicted}.

\input{table_worst_predicted.tex}

The last analysis we provide is the construction of the confusion matrices for all the participants. These are given in figures~\ref{fig:confusion}. Note the confusion matrices have been row normalized and the figures should be understood as the fraction of a given class that has been classified as such or such class.

\begin{figure}
	\begin{tabular}{ccccc}
		\includegraphics[width=0.18\columnwidth]{figs/confusion_matrix_spongebob.png}&
		\includegraphics[width=0.18\columnwidth]{figs/confusion_matrix_cestquilepatron.png}&
		\includegraphics[width=0.18\columnwidth]{figs/confusion_matrix_deepwhale.png}&
		\includegraphics[width=0.18\columnwidth]{figs/confusion_matrix_granderegazzoni.png}&
	\includegraphics[width=0.18\columnwidth]{conffigs/usion_matrix_TheBackpropagatedStubborns.png}\\
		SpongeBob & C'est qui l'patron & DeepWhalfigs/e & GrandeRegazzoni & The Backpropagated Stubborns\\
		\includegraphics[width=0.18\columnwidth]{figs/confusion_matrix_MetzDreamDeep.png}&
		\includegraphics[width=0.18\columnwidth]{figs/confusion_matrix_voleursDhuile.png}&
		\includegraphics[width=0.18\columnwidth]{figs/confusion_matrix_JMBmc.png}&
		\includegraphics[width=0.18\columnwidth]{figs/confusion_matrix_AG0D.png}&
	\includegraphics[width=0.18\columnwidth]{conffigs/usion_matrix_ResMetz.png}\\
		Metz Dream Deep & Voleurs d'Huile & JMBmcfigs/ & AG0D\\
		\includegraphics[width=0.18\columnwidth]{figs/confusion_matrix_CelestineFeuillat.png}&
		\includegraphics[width=0.18\columnwidth]{figs/confusion_matrix_LesJongleurs.png}&
		\includegraphics[width=0.18\columnwidth]{figs/confusion_matrix_DeepBeru'4s.png}&
		\includegraphics[width=0.18\columnwidth]{figs/confusion_matrix_TheWhaleSharks.png}&\\
		CelestineFeuillat & LesJongleurs & DeepBeru'4s & TheWhaleSharks & 
	\end{tabular}
	\caption{\label{fig:confusion} Confusion matrices of the best submissions of all the participants.}
\end{figure}

The confusion matrices allow to understand, when samples are misclassified, which are the most likely predicted classes.  Interestingly, the top ranking entries all share similar error patterns.

First, there is a tendency to misclassify as class 1 which is the detritus class, the majority class. Second, for the top ranking entries, when there is misclassification the misclassification is note widespread over all the classes but peaky on only few classes. Indeed, although this is hard to see from the plot, the confusion matrix is very sparse.

Looking at the misclassification, it reveals that some of them would have been hard to prevent and it reveals an issue in the challenge preparation. For example, looking at the samples provided in appendix, the class 33 "Nectophore Abylopsitetragona" seems pretty similar to the class 52 "Nectophore Diphyidae".

%TODO : check why the off diagonal elements class do not match the badly classified classes listed above. There seems to be issues for class 11 for example ..

%-----------------------------------------------------------------------------------
\appendix

\section*{The distribution of samples per class}

The table below indicates the number of samples per class for each fold. The list is ordered from the most frequent to the least frequent classes.

% python3 plot_stats.py /opt/Datasets/ZooScan/train /opt/Datasets/ZooScan/derived.csv
\input{stat_table.tex}

\pagebreak\newpage
\section*{Some examples for each class}

\input{./collage/examples.tex}

\end{document}

### 2022-05-06 Report/scripts/collage.py
collage and stats scripts

@@ -1,0 +1,53 @@
+#!/usr/bin/env python3
+
+import PIL
+import sys
+import pathlib
+import random
+import matplotlib.pyplot as plt
+import numpy as np
+
+
+def make_collage(traindir, outputdir, N=10):
+    dirs = [d for d in traindir.glob("*") if d.is_dir()]
+    outputdir.mkdir()
+    # Make the tex file to be included
+    texfile = open(outputdir / "examples.tex", "w")
+
+    # Let us sort the images by their class label
+    dirs = sorted(dirs, key=lambda d: d.parts[-1])
+
+    for d in dirs:
+        # Get some random samples for this class
+        print(f"Processing {d}")
+        files = list(d.glob("*.jpg"))
+        random.shuffle(files)
+
+        fig, axes = plt.subplots(1, N)
+        for ax, filename in zip(axes, files):
+            img = PIL.Image.open(filename, "r")
+            img = np.asarray(img)
+            ax.imshow(img, cmap="gray")
+            ax.axis("off")
+        classname = d.parts[-1]
+        plt.tight_layout()
+        # plt.suptitle(classname)
+        plt.savefig(outputdir / f"{classname}.jpg", bbox_inches="tight")
+        plt.close(fig)
+        local_path = outputdir.parts[-1] + "/" + f"{classname}.jpg"
+        classname = " ".join(classname.split("_")[1:])
+        texfile.write(
+            "\\begin{figure}[h]\n"
+            f"\\includegraphics[width=\\columnwidth]{{{local_path}}}"
+            f"\\caption{{Samples from the class {classname} }}\n"
+            "\\end{figure}\n"
+        )
+    texfile.close()
+
+
+if __name__ == "__main__":
+    if len(sys.argv) != 3:
+        print(f"Usage : {sys.argv[0]} root_traindir outputdir")
+        sys.exit(-1)
+
+    make_collage(pathlib.Path(sys.argv[1]), pathlib.Path(sys.argv[2]))

### 2022-05-06 Report/scripts/plot_stats.py

@@ -1,0 +1,71 @@
+# coding: utf-8
+import sys
+import os
+import glob
+import pandas as pd
+import matplotlib.pyplot as plt
+
+if len(sys.argv) != 3:
+    print(f"Usage {sys.argv[0]} traindir testcsv")
+    sys.exit(-1)
+
+traindir = sys.argv[1]
+testcsv = sys.argv[2]
+
+stats = pd.read_csv(testcsv, sep=",", header=1)
+stats.columns = ["filename", "class", "set"]
+
+private_test = stats[stats["set"] == "Private"]
+private_class_distrib = private_test["class"].value_counts()
+private_class_distrib = pd.DataFrame(
+    {"class": private_class_distrib.index, "Private test": private_class_distrib.values}
+)
+private_class_distrib = private_class_distrib.set_index("class")
+
+public_test = stats[stats["set"] == "Public"]
+public_class_distrib = public_test["class"].value_counts()
+public_class_distrib = pd.DataFrame(
+    {"class": public_class_distrib.index, "Public test": public_class_distrib.values}
+)
+public_class_distrib = public_class_distrib.set_index("class")
+print(public_class_distrib)
+
+# Collect the training set stats and class names
+class_names = {}
+
+trainset = []
+for path in os.scandir(traindir):
+    id_name = path.name.split("_")
+    class_names[int(id_name[0])] = " ".join(id_name[1:])
+    l = os.listdir(path)
+    trainset.append((int(id_name[0]), len(l)))
+trainset = pd.DataFrame(trainset, columns=["class", "Train"])
+trainset = trainset.set_index("class")
+
+df = public_class_distrib.join(private_class_distrib).join(trainset)
+print(df.cumsum())
+df.plot.bar(figsize=(15, 5))
+
+plt.xticks(horizontalalignment="center")
+plt.yscale("log")
+plt.ylabel("Number of samples")
+plt.tight_layout()
+# plt.show()
+plt.savefig("stats.pdf")
+
+# Generate the latex table
+df.reset_index()
+with open("stat_table.tex", "w") as f:
+    f.write("\\begin{tabular}{|c|r|c|c|c|}\n")
+    f.write("Index & Name & Train & Public & Private\\\\\n\hline")
+    for idx, (index, row) in enumerate(df.iterrows()):
+        f.write(
+            f"{index}  & {class_names[index]} & {row['Train']}& {row['Public test']}& {row['Private test']}\\\\\n"
[…]

### 2022-05-06 Report/collage/examples.tex
default collage of samples

\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/000_Candaciidae.jpg}\caption{Samples from the class Candaciidae }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/001_detritus.jpg}\caption{Samples from the class detritus }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/002_Calocalanus_pavo.jpg}\caption{Samples from the class Calocalanus pavo }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/003_larvae__Crustacea.jpg}\caption{Samples from the class larvae  Crustacea }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/004_Podon.jpg}\caption{Samples from the class Podon }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/005_Sapphirinidae.jpg}\caption{Samples from the class Sapphirinidae }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/006_Calanidae.jpg}\caption{Samples from the class Calanidae }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/007_zoea__Decapoda.jpg}\caption{Samples from the class zoea  Decapoda }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/008_Gammaridea.jpg}\caption{Samples from the class Gammaridea }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/009_Oikopleuridae.jpg}\caption{Samples from the class Oikopleuridae }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/010_Hyperiidea.jpg}\caption{Samples from the class Hyperiidea }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/011_zoea__Galatheidae.jpg}\caption{Samples from the class zoea  Galatheidae }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/012_nectophore__Physonectae.jpg}\caption{Samples from the class nectophore  Physonectae }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/013_Rhincalanidae.jpg}\caption{Samples from the class Rhincalanidae }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/014_Acantharea.jpg}\caption{Samples from the class Acantharea }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/015_Foraminifera.jpg}\caption{Samples from the class Foraminifera }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/016_nauplii__Crustacea.jpg}\caption{Samples from the class nauplii  Crustacea }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/017_gonophore__Diphyidae.jpg}\caption{Samples from the class gonophore  Diphyidae }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/018_metanauplii.jpg}\caption{Samples from the class metanauplii }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/019_megalopa.jpg}\caption{Samples from the class megalopa }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/020_Brachyura.jpg}\caption{Samples from the class Brachyura }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/021_tail__Chaetognatha.jpg}\caption{Samples from the class tail  Chaetognatha }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/022_Doliolida.jpg}\caption{Samples from the class Doliolida }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/023_Scyphozoa.jpg}\caption{Samples from the class Scyphozoa }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/024_Ctenophora.jpg}\caption{Samples from the class Ctenophora }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/025_Bivalvia__Mollusca.jpg}\caption{Samples from the class Bivalvia  Mollusca }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/026_ephyra.jpg}\caption{Samples from the class ephyra }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/027_Temoridae.jpg}\caption{Samples from the class Temoridae }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/028_scale.jpg}\caption{Samples from the class scale }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/029_Evadne.jpg}\caption{Samples from the class Evadne }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/030_Copilia.jpg}\caption{Samples from the class Copilia }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/031_Eucalanidae.jpg}\caption{Samples from the class Eucalanidae }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/032_Pyrosomatida.jpg}\caption{Samples from the class Pyrosomatida }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/033_nectophore__Abylopsis_tetragona.jpg}\caption{Samples from the class nectophore  Abylopsis tetragona }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/034_Actinopterygii.jpg}\caption{Samples from the class Actinopterygii }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/035_Creseidae.jpg}\caption{Samples from the class Creseidae }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/036_Calanoida.jpg}\caption{Samples from the class Calanoida }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/037_Decapoda.jpg}\caption{Samples from the class Decapoda }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/038_Obelia.jpg}\caption{Samples from the class Obelia }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/039_Noctiluca.jpg}\caption{Samples from the class Noctiluca }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/040_Spumellaria.jpg}\caption{Samples from the class Spumellaria }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/041_Chaetognatha.jpg}\caption{Samples from the class Chaetognatha }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/042_Annelida.jpg}\caption{Samples from the class Annelida }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/043_larvae__Annelida.jpg}\caption{Samples from the class larvae  Annelida }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/044_Rhopalonema.jpg}\caption{Samples from the class Rhopalonema }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/045_egg__other.jpg}\caption{Samples from the class egg  other }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/046_tail__Appendicularia.jpg}\caption{Samples from the class tail  Appendicularia }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/047_Euchirella.jpg}\caption{Samples from the class Euchirella }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/048_calyptopsis.jpg}\caption{Samples from the class calyptopsis }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/049_Haloptilus.jpg}\caption{Samples from the class Haloptilus }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/050_eudoxie__Diphyidae.jpg}\caption{Samples from the class eudoxie  Diphyidae }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/051_egg__Actinopterygii.jpg}\caption{Samples from the class egg  Actinopterygii }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/052_nectophore__Diphyidae.jpg}\caption{Samples from the class nectophore  Diphyidae }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/053_head.jpg}\caption{Samples from the class head }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/054_Penilia.jpg}\caption{Samples from the class Penilia }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/055_egg__Cavolinia_inflexa.jpg}\caption{Samples from the class egg  Cavolinia inflexa }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/056_Pontellidae.jpg}\caption{Samples from the class Pontellidae }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/057_Coscinodiscus.jpg}\caption{Samples from the class Coscinodiscus }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/058_Acartiidae.jpg}\caption{Samples from the class Acartiidae }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/059_Corycaeidae.jpg}\caption{Samples from the class Corycaeidae }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/060_artefact.jpg}\caption{Samples from the class artefact }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/061_cirrus.jpg}\caption{Samples from the class cirrus }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/062_Luciferidae.jpg}\caption{Samples from the class Luciferidae }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/063_Limacinidae.jpg}\caption{Samples from the class Limacinidae }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/064_cyphonaute.jpg}\caption{Samples from the class cyphonaute }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/065_part__Copepoda.jpg}\caption{Samples from the class part  Copepoda }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/066_Fritillariidae.jpg}\caption{Samples from the class Fritillariidae }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/067_Echinoidea.jpg}\caption{Samples from the class Echinoidea }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/068_Neoceratium.jpg}\caption{Samples from the class Neoceratium }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/069_Phaeodaria.jpg}\caption{Samples from the class Phaeodaria }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/070_Ostracoda.jpg}\caption{Samples from the class Ostracoda }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/071_Centropagidae.jpg}\caption{Samples from the class Centropagidae }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/072_Ophiuroidea.jpg}\caption{Samples from the class Ophiuroidea }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/073_nauplii__Cirripedia.jpg}\caption{Samples from the class nauplii  Cirripedia }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/074_Salpida.jpg}\caption{Samples from the class Salpida }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/075_Oithonidae.jpg}\caption{Samples from the class Oithonidae }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/076_eudoxie__Abylopsis_tetragona.jpg}\caption{Samples from the class eudoxie  Abylopsis tetragona }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/077_cypris.jpg}\caption{Samples from the class cypris }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/078_Oncaeidae.jpg}\caption{Samples from the class Oncaeidae }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/079_gonophore__Abylopsis_tetragona.jpg}\caption{Samples from the class gonophore  Abylopsis tetragona }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/080_Harpacticoida.jpg}\caption{Samples from the class Harpacticoida }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/081_Cavoliniidae.jpg}\caption{Samples from the class Cavoliniidae }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/082_Aglaura.jpg}\caption{Samples from the class Aglaura }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/083_Euchaetidae.jpg}\caption{Samples from the class Euchaetidae }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/084_Tomopteridae.jpg}\caption{Samples from the class Tomopteridae }
\end{figure}
\begin{figure}[h]
\includegraphics[width=\columnwidth]{collage/085_Limacidae.jpg}\caption{Samples from the class Limacidae }
\end{figure}

### 2022-05-06 Report/scripts/parse_scores.py
add the python script to parse the kaggle score page

@@ -1,0 +1,37 @@
+import pandas as pd
+import matplotlib.pyplot as plt
+import sys
+
+if len(sys.argv) != 2:
+    print(f"Usage : {sys.argv[0]} kagglescorepage.html")
+    sys.exit(-1)
+
+kaggle_html = sys.argv[1]
+
+num_participants = 13
+
+dfs = pd.read_html(kaggle_html)
+fig, ax = plt.subplots()
+for i in range(1, num_participants + 1):
+    dfs[i]["Date"] = pd.to_datetime(dfs[i]["Date"])
+    dfs[i]["Public"] = pd.to_numeric(dfs[i]["Public"], errors="coerce")
+    df = dfs[i].sort_values(by="Date", ascending=True)
+    df["Public"] = df["Public"].cummax()
+    df.plot(x="Date", y="Public", ax=ax, legend=False)
+ax.set_title("Scores through time on the public test set")
+ax.set_ylabel("Macro F1 score")
+ax.set_ylim(0.0, 1)
+plt.savefig("../figs/public-test.pdf", bbox_inches="tight")
+
+dfs = pd.read_html(kaggle_html)
+fig, ax = plt.subplots()
+for i in range(1, num_participants + 1):
+    dfs[i]["Date"] = pd.to_datetime(dfs[i]["Date"])
+    dfs[i]["Private"] = pd.to_numeric(dfs[i]["Private"], errors="coerce")
+    df = dfs[i].sort_values(by="Date", ascending=True)
+    df["Private"] = df["Private"].cummax()
+    df.plot(x="Date", y="Private", ax=ax, legend=False)
+ax.set_title("Scores through time on the public test set")
+ax.set_ylabel("Macro F1 score")
+ax.set_ylim(0.0, 1)
+plt.savefig("../figs/private-test.pdf", bbox_inches="tight")

### 2022-05-06 Report/scripts/plot_best_submissions.py
submissions F1 analysis script

@@ -1,0 +1,163 @@
+# coding: utf-8
+
+import sys
+import pathlib
+import pandas as pd
+import numpy as np
+import matplotlib.pyplot as plt
+import seaborn as sns
+
+if __name__ == "__main__":
+
+    if len(sys.argv) != 4:
+        print(f"Usage: {sys.argv[0]} gt_labels.csv csvdir class_names.csv")
+        print(f"e.g. python3 {sys.argv[0]} gt_all.csv participants_csv class_names.csv")
+        sys.exit(-1)
+
+    gt_labelfilename = sys.argv[1]
+    csvdir = pathlib.Path(sys.argv[2])
+    classnames = list(map(lambda x: x.rstrip(), open(sys.argv[3], "r").readlines()))
+    print(classnames)
+
+    df = pd.read_csv(
+        gt_labelfilename, index_col=0, usecols=["imgname", "label", "Usage"]
+    )
+
+    # Now we fill in the submissions of the participants
+    teams = []
+    for f in csvdir.glob("*.csv"):
+        team_name = f.name[:-4].split("_")[1]
+        submission_df = pd.read_csv(f, index_col=0)
+        submission_df = submission_df.rename(columns={"label": team_name})
+        df = pd.concat([df, submission_df], axis=1, join="inner")
+        teams.append(team_name)
+
+    print(f"I collected the stats about the following {len(teams)} teams : {teams}")
+
+    # If necessary, split on the board type (public or private)
+    # private_set = df[df["Usage"] == "Private"]
+    # public_set = df[df["Usage"] == "Public"]
+
+    # For info, count the number of class values per class
+    class_stats = df["label"].value_counts().to_frame("counts")
+    num_classes = len(class_stats)
+    print(f"Class counts for the {num_classes} classes : \n {class_stats}")
+
+    # Measure the class F1 scores for every teams and every class
+    for team in teams:
+        pred_labels = df[team]
+        true_labels = df["label"]
+        # for every class, we count the TP, FN, FP
+        # 1/F1 = 1/recall + 1/precision
+        # precision = TP / (TP + FP)
+        # recall = TP / (TP + FN)
+        f1s = {"index": [], team: []}
+        for cid in range(num_classes):
+            pred_labels_cid = (pred_labels == cid).to_numpy().astype("int")
+            true_labels_cid = (true_labels == cid).to_numpy().astype("int")
[…]

### 2022-05-06 Report/scripts/confusion_matrices.py
script to produce the confusion matrices

@@ -1,0 +1,62 @@
+# coding: utf-8
+
+import sys
+import pathlib
+import pandas as pd
+import matplotlib.pyplot as plt
+import seaborn as sns
+
+
+def confusion_matrix(num_classes, df):
+    crosstab = pd.crosstab(df["label"], df["predicted"], normalize="index")
+    # Some indices might be missing in the "predicted"
+    # because the predictor may never predict that class
+    indices = list(range(num_classes))
+    crosstab = crosstab.reindex(indices, axis=1).fillna(0)
+    return crosstab
+
+
+if __name__ == "__main__":
+
+    if len(sys.argv) != 3:
+        print(f"Usage: {sys.argv[0]} gt_labels.csv csvdir")
+        print(f"e.g. python3 {sys.argv[0]} gt_all.csv participants_csv")
+        sys.exit(-1)
+
+    # Merge the csvs and the ground truth
+    gt_labelfilename = sys.argv[1]
+    csvdir = pathlib.Path(sys.argv[2])
+
+    df = pd.read_csv(
+        gt_labelfilename, index_col=0, usecols=["imgname", "label", "Usage"]
+    )
+
+    # Now we fill in the submissions of the participants
+    teams = []
+    for f in csvdir.glob("*.csv"):
+        team_name = f.name[:-4].split("_")[1]
+        submission_df = pd.read_csv(f, index_col=0)
+        submission_df = submission_df.rename(columns={"label": team_name})
+        df = pd.concat([df, submission_df], axis=1, join="inner")
+        teams.append(team_name)
+
+    class_stats = df["label"].value_counts().to_frame("counts")
+    num_classes = len(class_stats)
+
+    # Compute the confusion matrices, for each participant
+    for team in teams:
+        print(f"Evaluating the confusion matrix for the team {team}")
+        subdf = df[["label", team]].rename(columns={team: "predicted"})
+        print(subdf.head())
+        confusion_mtx = confusion_matrix(num_classes, subdf)
+        print(confusion_mtx)
+        f = plt.figure(figsize=(15, 15))
+        sns.heatmap(confusion_mtx, cmap="Blues", vmin=0, vmax=1)
+        # plt.imshow(confusion_mtx, vmin=0, vmax=1, aspect="equal")
+
+        ax = plt.gca()
+        ax.set_xlim(0, num_classes)
+        ax.set_ylim(num_classes, 0)
+        plt.grid()
[…]

### 2022-05-06 Report/paper.tex
fix path, show sizes
fix paper fig/table paths

@@
 	\begin{center}
-		% \includegraphics[width=0.3\columnwidth]{figs/size_train.png}
-		% \includegraphics[width=0.3\columnwidth]{figs/sizes_test.pdf}
-		% \includegraphics[width=0.3\columnwidth]{figs/size_orig.png}
+		\includegraphics[width=0.3\columnwidth]{figs/size_train.png}
+		\includegraphics[width=0.3\columnwidth]{figs/sizes_test.png}
+		\includegraphics[width=0.3\columnwidth]{figs/size_orig.png}
 	\end{center}
@@
 
-\input{table_best_predicted.tex}
+\input{./figs/table_best_predicted.tex}
 
@@
 
-\input{table_worst_predicted.tex}
+\input{./figs/table_worst_predicted.tex}
 
@@
 		\includegraphics[width=0.18\columnwidth]{figs/confusion_matrix_granderegazzoni.png}&
-	\includegraphics[width=0.18\columnwidth]{conffigs/usion_matrix_TheBackpropagatedStubborns.png}\\
+        \includegraphics[width=0.18\columnwidth]{figs/confusion_matrix_TheBackpropagatedStubborns.png}\\
 		SpongeBob & C'est qui l'patron & DeepWhalfigs/e & GrandeRegazzoni & The Backpropagated Stubborns\\
@@
 		\includegraphics[width=0.18\columnwidth]{figs/confusion_matrix_AG0D.png}&
-	\includegraphics[width=0.18\columnwidth]{conffigs/usion_matrix_ResMetz.png}\\
+        \includegraphics[width=0.18\columnwidth]{figs/confusion_matrix_ResMetz.png}\\
 		Metz Dream Deep & Voleurs d'Huile & JMBmcfigs/ & AG0D\\
@@
 % python3 plot_stats.py /opt/Datasets/ZooScan/train /opt/Datasets/ZooScan/derived.csv
-\input{stat_table.tex}
+\input{./figs/stat_table.tex}
 

### 2022-05-06 Report/paper.tex
page break before the appendix
fix confusion matrix legend

@@
         \includegraphics[width=0.18\columnwidth]{figs/confusion_matrix_TheBackpropagatedStubborns.png}\\
-		SpongeBob & C'est qui l'patron & DeepWhalfigs/e & GrandeRegazzoni & The Backpropagated Stubborns\\
+		SpongeBob & C'est qui l'patron & DeepWhale & GrandeRegazzoni & The Backpropagated Stubborns\\
 		\includegraphics[width=0.18\columnwidth]{figs/confusion_matrix_MetzDreamDeep.png}&
@@
         \includegraphics[width=0.18\columnwidth]{figs/confusion_matrix_ResMetz.png}\\
-		Metz Dream Deep & Voleurs d'Huile & JMBmcfigs/ & AG0D\\
+        Metz Dream Deep & Voleurs d'Huile & JMBmc & AG0D & ResMetz \\
 		\includegraphics[width=0.18\columnwidth]{figs/confusion_matrix_CelestineFeuillat.png}&
@@
 %-----------------------------------------------------------------------------------
+\clearpage
+
 \appendix

### 2022-05-06 SampleCode/data.py
base code

@@ -1,0 +1,579 @@
+# coding: utf-8
+
+# Standard imports
+import logging
+import pathlib
+import os
+import shutil
+import argparse
+import tqdm
+import glob
+import random
+
+# External imports
+import torchvision
+import torchvision.transforms as transforms
+import torch
+from torch.distributions import Beta
+import numpy as np
+import matplotlib.pyplot as plt
+from typing import Any, Callable, List, Optional, Tuple
+from PIL import ImageOps
+import cv2
+import albumentations as A
+from albumentations.core.transforms_interface import ImageOnlyTransform
+from albumentations.pytorch import ToTensorV2
+
+__TRAIN_URL = "http://infomob.metz.supelec.fr/fix/ChallengeDeep/train.tar.gz"
+__TEST_URL = "http://infomob.metz.supelec.fr/fix/ChallengeDeep/test.tar.gz"
+
+
+class Resize:
+    """
+    Resize the image to a square image of target_size x target_size
+    preserving the aspect ratio
+    """
+
+    def __init__(self, target_size: int):
+        self.target_size = target_size
+
+    def __call__(self, pic):
+        max_size = max(pic.width, pic.height)
+        pic = ImageOps.grayscale(pic)
+        # The image is centered with the padding
+        pic = ImageOps.pad(pic, (max_size, max_size), color=(255,))
+        pic = ImageOps.fit(pic, (self.target_size, self.target_size))
+        return pic
+
+    def __repr__(self):
+        return self.__class__.__name__ + f"(self.target_size)"
+
+
+class MyDataset(torch.utils.data.Dataset):
+    """
+    Wrapper class around the ImageFolder dataset to add the static features as
+    well
+    """
+
+    def __init__(
+        self,
+        root: str,
+        transform: Optional[Callable] = None,
+        target_transform: Optional[Callable] = None,
+    ) -> None:
+        root_path = pathlib.Path(root)
+        if not root_path.exists():
+            # TODO: download the data
+            # and extract the archive
+            print(f"Path {root_path} does not exist")
+        self.imgdataset = torchvision.datasets.ImageFolder(
+            root, transform, target_transform
+        )
+
+    def __getitem__(self, index: int) -> Any:
+        return self.imgdataset[index]
+
+    @property
+    def samples(self):
+        return self.imgdataset.samples
+
+    @property
[…]

### 2022-05-06 SampleCode/job.py

@@ -1,0 +1,110 @@
+#!/usr/bin/env python3
+
+"""
+Script used to run experiments on a slurm cluster using the commit-id to guarantee the executed code version
+"""
+
+import os
+import subprocess
+
+
+def makejob(commit_id, nruns, partition, walltime, normalize, params):
+    paramsstr = " ".join([f"--{name} {value}" for name, value in params.items()])
+    if normalize:
+        paramsstr += " --normalize "
+    return f"""#!/bin/bash
+
+#SBATCH --job-name=challenge-{params['model']}
+#SBATCH --nodes=1
+#SBATCH --partition={partition}
+#SBATCH --time={walltime}
+#SBATCH --output=logslurms/slurm-%A_%a.out
+#SBATCH --error=logslurms/slurm-%A_%a.err
+#SBATCH --array=0-{nruns-1}
+
+current_dir=`pwd`
+
+# Fix env variables as bashrc and profile are not loaded
+export LOCAL=$HOME/.local
+export PATH=$PATH:$LOCAL/bin
+
+echo "Session " {params['model']}_${{SLURM_ARRAY_JOB_ID}}_${{SLURM_ARRAY_TASK_ID}}
+date
+
+echo "Running on $(hostname)"
+# env
+
+echo "Copying the source directory and data"
+date
+mkdir $TMPDIR/challenge
+cd ..
+rsync -r .. $TMPDIR/ --exclude 'SampleCode/logslurms' --exclude 'DatasetPreparation' --exclude Report 
+
+cd $TMPDIR/SampleCode/
+git checkout {commit_id}
+
+echo ""
+echo "Virtual env"
+
+python3 -m pip install virtualenv --user
+virtualenv -p python3 venv
+source venv/bin/activate
+python -m pip install -r requirements.txt
+
+
+
+echo ""
+echo "Training"
+date
+
+python3 main.py  --datadir ./ChallengeDeep/training {paramsstr} --logname {params['model']}_${{SLURM_ARRAY_JOB_ID}}_${{SLURM_ARRAY_TASK_ID}} --commit_id '{commit_id}' --logdir ${{current_dir}}/logs train
+
+if [[ $? != 0 ]]; then
+    exit -1
+fi
+date
+"""
+
+
+def submit_job(job):
+    with open("job.sbatch", "w") as fp:
+        fp.write(job)
+    os.system("sbatch job.sbatch")
+
+
+# Ensure all the modified files have been staged and commited
+result = int(
[…]

### 2022-05-06 SampleCode/main.py

@@ -1,0 +1,469 @@
+#!/usr/bin/env python3
+# coding: utf-8
+
+# Standard modules
+import logging
+import argparse
+import pathlib
+import os
+import shutil
+import itertools
+import sys
+import math
+from typing import Optional
+# External modules
+import tqdm
+import deepcs
+import deepcs.training
+import deepcs.testing
+import deepcs.metrics
+import deepcs.display
+import deepcs.fileutils
+import torch
+import torch.nn as nn
+import torch.nn.functional as F
+import torch.optim
+from torch.utils.tensorboard import SummaryWriter
+import torchvision.transforms as transforms
+import pandas as pd
+from sklearn.metrics import f1_score
+import matplotlib
+matplotlib.use('Agg')
+import matplotlib.pyplot as plt
+import numpy as np
+import cv2
+from PIL import ImageOps
+import albumentations as A
+from albumentations.pytorch import ToTensorV2
+# Local modules
+import data
+import models
+
+def accuracy(probabilities, targets):
+    """
+    Computes the accuracy. Works with either PackedSequence or Tensor
+    """
+    with torch.no_grad():
+        if isinstance(probabilities, torch.nn.utils.rnn.PackedSequence):
+            probs = probabilities.data
+        else:
+            probs = probabilities
+        if isinstance(targets, torch.nn.utils.rnn.PackedSequence):
+            targ = targets.data
+        else:
+            targ = targets
+        if len(targ.shape) == 2:
+            # For the LabelSmooth case
+            targ = targ.argmax(axis=-1) 
+        return (probs.argmax(axis=-1) == targ).double().mean()
+
+def f1_metric(model, loader, device):
+    with torch.no_grad():
+        model.eval()
+        pred_labels, true_labels = [], []
+        for (inputs, targets) in loader:
+            inputs, targets = inputs.to(device), targets.to(device)
+            outputs = model(inputs)
+
+            predicted_labels = outputs.argmax(axis=-1).detach().tolist()
+            pred_labels.extend(predicted_labels)
+            if len(targets.shape) == 2:
+                # For the LabelSmooth case
+                targets = targets.argmax(axis=-1) 
+            true_labels.extend(targets.detach().tolist())
+        macro_F1 = f1_score(true_labels, pred_labels, average='macro')
+        class_F1 = f1_score(true_labels, pred_labels, average=None)
+    return macro_F1, class_F1
+
+
+
+def compute_class_weights(n_samples_per_class):
[…]

### 2022-05-06 SampleCode/models.py

@@ -1,0 +1,285 @@
+#!/usr/bin/env python3
+
+import torch
+import torch.nn as nn
+import torchvision.models
+import torchvision.models.resnet as resnet
+import timm
+import timm.models.layers
+
+
+class LinearNet(nn.Module):
+    def __init__(self, img_size, num_classes):
+        super(LinearNet, self).__init__()
+        dim = img_size[0] * img_size[1]
+        self.model = nn.Linear(dim, num_classes)
+
+    def forward(self, x):
+        x = x.view(x.size()[0], -1)
+        return self.model(x)
+
+
+def conv_bn_relu(in_channels, out_channels, ks):
+    return [
+        nn.Conv2d(
+            in_channels,
+            out_channels,
+            kernel_size=ks,
+            stride=1,
+            padding=int((ks - 1) / 2),
+            bias=True,
+        ),
+        nn.BatchNorm2d(out_channels),
+        nn.ReLU(inplace=True),
+    ]
+
+
+def conv_block(fin, fout, ks):
+    return [
+        *conv_bn_relu(fin, fout, ks),
+        *conv_bn_relu(fout, fout, ks),
+        *conv_bn_relu(fout, fout, ks),
+    ]
+
+
+class SimpleCNNRF(nn.Module):
+    def __init__(self, img_size, num_classes):
+        super().__init__()
+        self.features = nn.Sequential(
+            *conv_block(1, 32, 5),
+            nn.MaxPool2d(kernel_size=2),
+            *conv_block(32, 64, 5),
+            nn.MaxPool2d(kernel_size=2),
+            *conv_block(64, 128, 5),
+            nn.MaxPool2d(kernel_size=2),
+            *conv_block(128, 128, 5),
+            nn.MaxPool2d(kernel_size=2),
+            *conv_block(128, 128, 5),
+        )
+        # Probe the output shape
+        probe_tensor = torch.zeros((1, 1, img_size[0], img_size[1]))
+        out = self.features(probe_tensor)
+        dim = out.view(-1).shape[0]
+
+        # Make the classifier head
+        self.classifier = nn.Sequential(nn.Dropout2d(0.5), nn.Linear(dim, num_classes))
+
+    def forward(self, x):
+        self.feature_maps = []
+        for l in self.features:
+            x = l(x)
+            self.feature_maps.append(x)
+        x = x.view(x.size()[0], -1)
+        return self.classifier(x)
+
+
[…]

### 2022-05-06 SampleCode/preprocess.py

@@ -1,0 +1,68 @@
+#!/usr/bin/env python3
+# coding: utf-8
+
+"""
+Preprocess the data and save the resized, cropped, etc.. images
+
+Takes a directory as dir/000_xxx/i.jpg , ...
+loads the images, resize them , pad, crop , etc..
+
+And saves the resulting images into dir/train/000_xxx/j.jpg
+ and                                dir/valid/000_xxx/k.jpg
+"""
+
+# Standard modules
+import logging 
+import argparse
+import pathlib
+import os
+import shutil
+# External modules
+import tqdm 
+# Local modules
+import data
+
+def process_dataset(dataset, outputdir):
+    img_idx = {}
+    logging.info(f"Processsing {len(dataset)} objects in {outputdir}")
+    for X, y in tqdm.tqdm(dataset):
+        if y not in img_idx:
+            img_idx[y] = 0
+        output_filedir = outputdir / f"{y:03d}" 
+
+        if not output_filedir.exists():
+            os.makedirs(output_filedir)
+        if img_idx[y] >= 1e7:
+            raise ValueError("Our image filename format expects a number of images <= 1 000 000")
+        output_filepath = output_filedir / f"{img_idx[y]:06d}.jpg"
+        img_idx[y] += 1
+        X.save(output_filepath)
+
+
+def main(args):
+    logging.info("Loading the images")
+    train_dataset, valid_dataset = data.load_dataset(args.datadir,
+                                                transform=data.Resize(data.__default_size[0]),
+                                                val_ratio=args.val_ratio)
+    process_dataset(train_dataset, args.outputdir / "train")
+    logging.info("Done")
+    process_dataset(valid_dataset, args.outputdir / "valid")
+    logging.info("Done")
+
+
+if __name__ == '__main__':
+    logging.basicConfig(level=logging.INFO, format='%(message)s')
+    parser = argparse.ArgumentParser()
+    parser.add_argument('--datadir', type=pathlib.Path, required=True)
+    parser.add_argument('--outputdir', type=pathlib.Path, required=True)
+    parser.add_argument('--num_workers', type=int, default=4)
+    parser.add_argument('--batch_size', type=int, default=64)
+    parser.add_argument('--val_ratio', type=float, default=0.1)
+
+    args = parser.parse_args()
+    if args.outputdir.exists():
+        logging.info("The datadir exists, we use it")
+        # logging.info(f"Removing {outputdir}")
+        # shutil.rmtree(outputdir)
+    else:
+        main(args)

### 2022-05-06 SampleCode/preprocess.py
blake refactor

@@ -12,15 +12,18 @@ And saves the resulting images into dir/train/000_xxx/j.jpg
 """
 
 # Standard modules
-import logging 
+import logging
 import argparse
 import pathlib
 import os
 import shutil
+
 # External modules
-import tqdm 
+import tqdm
+
 # Local modules
 import data
+
 
 def process_dataset(dataset, outputdir):
     img_idx = {}
@@ -28,12 +31,14 @@ def process_dataset(dataset, outputdir):
     for X, y in tqdm.tqdm(dataset):
         if y not in img_idx:
             img_idx[y] = 0
-        output_filedir = outputdir / f"{y:03d}" 
+        output_filedir = outputdir / f"{y:03d}"
 
         if not output_filedir.exists():
             os.makedirs(output_filedir)
         if img_idx[y] >= 1e7:
-            raise ValueError("Our image filename format expects a number of images <= 1 000 000")
+            raise ValueError(
+                "Our image filename format expects a number of images <= 1 000 000"
+            )
         output_filepath = output_filedir / f"{img_idx[y]:06d}.jpg"
         img_idx[y] += 1
         X.save(output_filepath)
@@ -41,23 +46,25 @@ def process_dataset(dataset, outputdir):
 
 def main(args):
     logging.info("Loading the images")
-    train_dataset, valid_dataset = data.load_dataset(args.datadir,
-                                                transform=data.Resize(data.__default_size[0]),
-                                                val_ratio=args.val_ratio)
+    train_dataset, valid_dataset = data.load_dataset(
+        args.datadir,
+        transform=data.Resize(data.__default_size[0]),
+        val_ratio=args.val_ratio,
+    )
     process_dataset(train_dataset, args.outputdir / "train")
     logging.info("Done")
     process_dataset(valid_dataset, args.outputdir / "valid")
     logging.info("Done")
 
 
-if __name__ == '__main__':
-    logging.basicConfig(level=logging.INFO, format='%(message)s')
+if __name__ == "__main__":
+    logging.basicConfig(level=logging.INFO, format="%(message)s")
     parser = argparse.ArgumentParser()
-    parser.add_argument('--datadir', type=pathlib.Path, required=True)
-    parser.add_argument('--outputdir', type=pathlib.Path, required=True)
-    parser.add_argument('--num_workers', type=int, default=4)
-    parser.add_argument('--batch_size', type=int, default=64)
[…]

### 2022-05-06 SampleCode/preprocess.py
better doc the image field size; change the ratio to 0.95/0.05

@@ -16,13 +16,14 @@ import logging
 import argparse
 import pathlib
 import os
-import shutil
 
 # External modules
 import tqdm
 
 # Local modules
 import data
+
+IMAGE_IDX_FILEWIDTH = 6
 
 
 def process_dataset(dataset, outputdir):
@@ -35,10 +36,15 @@ def process_dataset(dataset, outputdir):
 
         if not output_filedir.exists():
             os.makedirs(output_filedir)
-        if img_idx[y] >= 1e7:
+        if len(str(img_idx[y])) >= IMAGE_IDX_FILEWIDTH:
             raise ValueError(
-                "Our image filename format expects a number of images <= 1 000 000"
+                "Our image filename format expects a number of images <= 1"
+                + "0" * IMAGE_IDX_FILEWIDTH
             )
+        # If the following assert is not true
+        # we need to adapt the f-string format below
+        # do not know how to that with a variable
+        assert IMAGE_IDX_FILEWIDTH == 6
         output_filepath = output_filedir / f"{img_idx[y]:06d}.jpg"
         img_idx[y] += 1
         X.save(output_filepath)
@@ -64,12 +70,10 @@ if __name__ == "__main__":
     parser.add_argument("--outputdir", type=pathlib.Path, required=True)
     parser.add_argument("--num_workers", type=int, default=4)
     parser.add_argument("--batch_size", type=int, default=64)
-    parser.add_argument("--val_ratio", type=float, default=0.1)
+    parser.add_argument("--val_ratio", type=float, default=0.05)
 
     args = parser.parse_args()
     if args.outputdir.exists():
-        logging.info("The datadir exists, we use it")
-        # logging.info(f"Removing {outputdir}")
-        # shutil.rmtree(outputdir)
+        logging.info("The datadir exists, nothing to be done")
     else:
         main(args)

### 2022-05-06 SampleCode/data.py
licensing

@@ -1,13 +1,28 @@
 # coding: utf-8
 
+"""
+    This script belongs to the plankton classification sample code
+    Copyright (C) 2022 Jeremy Fix
+
+    This program is free software: you can redistribute it and/or modify
+    it under the terms of the GNU General Public License as published by
+    the Free Software Foundation, either version 3 of the License, or
+    (at your option) any later version.
+
+    This program is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+    GNU General Public License for more details.
+
+    You should have received a copy of the GNU General Public License
+    along with this program.  If not, see <https://www.gnu.org/licenses/>.
+"""
 # Standard imports
 import logging
 import pathlib
 import os
-import shutil
 import argparse
 import tqdm
-import glob
 import random
 
 # External imports
@@ -17,7 +32,7 @@ import torch
 from torch.distributions import Beta
 import numpy as np
 import matplotlib.pyplot as plt
-from typing import Any, Callable, List, Optional, Tuple
+from typing import Any, Callable, Optional
 from PIL import ImageOps
 import cv2
 import albumentations as A

### 2022-05-06 SampleCode/job.py

@@ -1,4 +1,21 @@
 #!/usr/bin/env python3
+"""
+    This script belongs to the plankton classification sample code
+    Copyright (C) 2022 Jeremy Fix
+
+    This program is free software: you can redistribute it and/or modify
+    it under the terms of the GNU General Public License as published by
+    the Free Software Foundation, either version 3 of the License, or
+    (at your option) any later version.
+
+    This program is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+    GNU General Public License for more details.
+
+    You should have received a copy of the GNU General Public License
+    along with this program.  If not, see <https://www.gnu.org/licenses/>.
+"""
 
 """
 Script used to run experiments on a slurm cluster using the commit-id to guarantee the executed code version

### 2022-05-06 SampleCode/main.py

@@ -1,5 +1,22 @@
 #!/usr/bin/env python3
 # coding: utf-8
+"""
+    This script belongs to the plankton classification sample code
+    Copyright (C) 2022 Jeremy Fix
+
+    This program is free software: you can redistribute it and/or modify
+    it under the terms of the GNU General Public License as published by
+    the Free Software Foundation, either version 3 of the License, or
+    (at your option) any later version.
+
+    This program is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+    GNU General Public License for more details.
+
+    You should have received a copy of the GNU General Public License
+    along with this program.  If not, see <https://www.gnu.org/licenses/>.
+"""
 
 # Standard modules
 import logging

### 2022-05-06 SampleCode/models.py

@@ -1,4 +1,21 @@
 #!/usr/bin/env python3
+"""
+    This script belongs to the plankton classification sample code
+    Copyright (C) 2022 Jeremy Fix
+
+    This program is free software: you can redistribute it and/or modify
+    it under the terms of the GNU General Public License as published by
+    the Free Software Foundation, either version 3 of the License, or
+    (at your option) any later version.
+
+    This program is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+    GNU General Public License for more details.
+
+    You should have received a copy of the GNU General Public License
+    along with this program.  If not, see <https://www.gnu.org/licenses/>.
+"""
 
 import torch
 import torch.nn as nn

### 2022-05-06 SampleCode/preprocess.py

@@ -1,5 +1,22 @@
 #!/usr/bin/env python3
 # coding: utf-8
+"""
+    This script belongs to the plankton classification sample code
+    Copyright (C) 2022 Jeremy Fix
+
+    This program is free software: you can redistribute it and/or modify
+    it under the terms of the GNU General Public License as published by
+    the Free Software Foundation, either version 3 of the License, or
+    (at your option) any later version.
+
+    This program is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+    GNU General Public License for more details.
+
+    You should have received a copy of the GNU General Public License
+    along with this program.  If not, see <https://www.gnu.org/licenses/>.
+"""
 
 """
 Preprocess the data and save the resized, cropped, etc.. images

### 2022-05-06 DatasetPreparation/prepare_data.py
licensing GPL

@@ -1,5 +1,23 @@
 #!/usr/bin/env python3
 # coding: utf-8
+"""
+    This script belongs to the plankton classification sample code
+    It prepares the dataset for the challenge
+    Copyright (C) 2022 Jeremy Fix
+
+    This program is free software: you can redistribute it and/or modify
+    it under the terms of the GNU General Public License as published by
+    the Free Software Foundation, either version 3 of the License, or
+    (at your option) any later version.
+
+    This program is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+    GNU General Public License for more details.
+
+    You should have received a copy of the GNU General Public License
+    along with this program.  If not, see <https://www.gnu.org/licenses/>.
+"""
 
 # Standard modules
 import argparse

### 2022-05-06 DatasetPreparation/prepare_data.py
licensing

@@ -101,6 +101,15 @@ def create_image(source_path, target_path, maxsize, rotation):
 
 
 if __name__ == "__main__":
+
+    license = """
+    prepare_data.py  Copyright (C) 2022  Jeremy Fix
+    This program comes with ABSOLUTELY NO WARRANTY;
+    This is free software, and you are welcome to redistribute it
+    under certain conditions;
+    """
+    print(license)
+
     logging.basicConfig(level=logging.INFO, format="%(message)s")
 
     parser = argparse.ArgumentParser()

### 2022-05-06 SampleCode/main.py

@@ -458,6 +458,14 @@ def test(args):
 
 
 if __name__ == "__main__":
+    license = """
+    prepare_data.py  Copyright (C) 2022  Jeremy Fix
+    This program comes with ABSOLUTELY NO WARRANTY;
+    This is free software, and you are welcome to redistribute it
+    under certain conditions;
+    """
+    print(license)
+
     logging.basicConfig(level=logging.INFO, format="%(message)s")
     parser = argparse.ArgumentParser()
     parser.add_argument("command", choices=["train", "test", "stats"])

### 2022-05-06 SampleCode/data.py
fix f string and rm useless variables

@@ -61,7 +61,7 @@ class Resize:
         return pic
 
     def __repr__(self):
-        return self.__class__.__name__ + f"(self.target_size)"
+        return self.__class__.__name__ + f"{self.target_size}"
 
 
 class MyDataset(torch.utils.data.Dataset):
@@ -377,7 +377,7 @@ def test_dataloader(args):
         num_workers=4,
         pin_memory=False,
     )
-    test_loader = load_test_data(
+    _ = load_test_data(
         pathlib.Path(args.datadir) / ".." / "pubtest",
         transform=__default_transform,
         batch_size=64,
@@ -404,7 +404,6 @@ class AugmentedDataset(torch.utils.data.Dataset):
     def __getitem__(self, idx):
         img, label = self.dataset[idx]
         img = np.array(img)
-        transf = self.transform(image=img)
         return self.transform(image=img)["image"], label
 
     @property
@@ -455,8 +454,6 @@ class KeepChannel(ImageOnlyTransform):
 
 
 def test_augmentations(args):
-    num_workers = 4
-    pin_memory = True
     num_imgs = 36
     transform = A.Compose(
         [
@@ -508,8 +505,6 @@ def test_augmented_dataloader(args):
             ToTensorV2(),
         ]
     )
-
-    use_cuda = torch.cuda.is_available()
     dataset = torchvision.datasets.ImageFolder(args.datadir)
     dataset = AugmentedDataset(dataset, transforms)
     for i in range(10):
@@ -523,7 +518,6 @@ def test_augmented_dataloader(args):
 def test_sampler(args):
     transforms = A.Compose([ToTensorV2()])
 
-    use_cuda = torch.cuda.is_available()
     dataset = torchvision.datasets.ImageFolder(args.datadir)
     dataset = AugmentedDataset(dataset, transforms)
     batch_size = 32

### 2022-05-06 SampleCode/main.py
refactored, documented utils

@@ -25,7 +25,6 @@ import pathlib
 import os
 import sys
 import math
-from typing import Optional
 
 # External modules
 import tqdm
@@ -42,12 +41,6 @@ import torch.optim
 from torch.utils.tensorboard import SummaryWriter
 import torchvision.transforms as transforms
 import pandas as pd
-from sklearn.metrics import f1_score
-import matplotlib
-
-matplotlib.use("Agg")
-import matplotlib.pyplot as plt
-import numpy as np
 import cv2
 from PIL import ImageOps
 import albumentations as A
@@ -56,88 +49,7 @@ from albumentations.pytorch import ToTensorV2
 # Local modules
 import data
 import models
-
-
-def accuracy(probabilities, targets):
-    """
-    Computes the accuracy. Works with either PackedSequence or Tensor
-    """
-    with torch.no_grad():
-        if isinstance(probabilities, torch.nn.utils.rnn.PackedSequence):
-            probs = probabilities.data
-        else:
-            probs = probabilities
-        if isinstance(targets, torch.nn.utils.rnn.PackedSequence):
-            targ = targets.data
-        else:
-            targ = targets
-        if len(targ.shape) == 2:
-            # For the LabelSmooth case
-            targ = targ.argmax(axis=-1)
-        return (probs.argmax(axis=-1) == targ).double().mean()
-
-
-def f1_metric(model, loader, device):
-    with torch.no_grad():
-        model.eval()
-        pred_labels, true_labels = [], []
-        for (inputs, targets) in loader:
-            inputs, targets = inputs.to(device), targets.to(device)
-            outputs = model(inputs)
-
-            predicted_labels = outputs.argmax(axis=-1).detach().tolist()
-            pred_labels.extend(predicted_labels)
-            if len(targets.shape) == 2:
-                # For the LabelSmooth case
-                targets = targets.argmax(axis=-1)
-            true_labels.extend(targets.detach().tolist())
-        macro_F1 = f1_score(true_labels, pred_labels, average="macro")
-        class_F1 = f1_score(true_labels, pred_labels, average=None)
-    return macro_F1, class_F1
-
-
-def compute_class_weights(n_samples_per_class):
-    """
-    n_samples_per_class:dict[int] -> int
-    """
-    # example count bounds for the training set
[…]

### 2022-05-06 SampleCode/utils.py

@@ -1,0 +1,170 @@
+# coding: utf-8
+"""
+    This script belongs to the plankton classification sample code
+    Copyright (C) 2022 Jeremy Fix
+
+    This program is free software: you can redistribute it and/or modify
+    it under the terms of the GNU General Public License as published by
+    the Free Software Foundation, either version 3 of the License, or
+    (at your option) any later version.
+
+    This program is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+    GNU General Public License for more details.
+
+    You should have received a copy of the GNU General Public License
+    along with this program.  If not, see <https://www.gnu.org/licenses/>.
+"""
+from typing import Optional
+import logging
+import torch
+import torch.data
+from sklearn.metrics import f1_score
+import matplotlib.pyplot as plt
+import numpy as np
+import tqdm
+
+
+def accuracy(probabilities: torch.Tensor, targets: torch.Tensor):
+    """Computes the accuracy. Works with either PackedSequence or Tensor
+
+    Args:
+        probabiliities (torch.Tensor (B, C)) : The predicted probabilities
+        targets (torch.Tensor (B,) ): The targets
+
+    Returns:
+        double: the accuracy
+    """
+    with torch.no_grad():
+        if len(targets.shape) == 2:
+            # For the LabelSmooth case
+            targets = targets.argmax(axis=-1)
+        return (probabilities.argmax(axis=-1) == targets).double().mean()
+
+
+def f1_metric(
+    model: torch.nn.Module, loader: torch.data.Dataloader, device: torch.device
+):
+    """Computes the F1 score of a model over datas provided by a dataloader
+
+    Args:
+        model (nn.Module): The module through which to forward propagate
+                           the inputs
+        loader (torch.data.Dataloader): a mappable dataloader
+        device (torch.device): The device on which to compute
+
+    Returns:
+        macro_F1 (double) : the macro average F1 score
+        class_F1 (double): the F1 score per class
+    """
+    with torch.no_grad():
+        pred_labels, true_labels = [], []
+        for (inputs, targets) in loader:
+            inputs, targets = inputs.to(device), targets.to(device)
+            outputs = model(inputs)
+
[…]

### 2022-05-06 SampleCode/main.py
fix and doc focal loss

@@ -52,30 +52,6 @@ import models
 import utils
 
 
-class FocalLoss(nn.Module):
-    def __init__(self, weights, gamma):
-        super(FocalLoss, self).__init__()
-        self.weights = weights
-        self.gamma = gamma
-        self.eps = 1e-10
-
-    def forward(self, predictions, target):
-        probs = F.softmax(predictions, dim=1) + self.eps
-        target_one_hot = utils.one_hot(
-            target,
-            num_classes=predictions.shape[1],
-            device=predictions.device,
-            dtype=predictions.dtype,
-        )
-
-        weight = torch.pow(1.0 - probs, self.gamma)
-        focal = -weight * torch.log(probs)
-        loss_tmp = torch.sum(target_one_hot * focal, dim=1)
-        loss = loss_tmp.mean()
-
-        return loss
-
-
 def stats(args):
     print("Stats")
 
@@ -191,7 +167,7 @@ def train(args):
     if args.loss == "BCE":
         loss = bce_loss
     else:
-        loss = utils.FocalLoss(class_weights, gamma=0.5)
+        loss = utils.FocalLoss(gamma=0.5)
 
     # Make the optimizer
     optimizer = torch.optim.Adam(

### 2022-05-06 SampleCode/utils.py

@@ -20,6 +20,8 @@ from typing import Optional
 import logging
 import torch
 import torch.data
+import torch.nn as nn
+import torch.nn.functional as F
 from sklearn.metrics import f1_score
 import matplotlib.pyplot as plt
 import numpy as np
@@ -168,3 +170,37 @@ def plot_confusion_matrix(confusion_matrix: np.array):
     plt.ylabel("Predicted")
     plt.xlabel("True")
     return fig
+
+
+class FocalLoss(nn.Module):
+    """
+    Implementation of the Focal Loss :math:`\frac{1}{N} \sum_i -(1-p_{y_i} + \epsilon)^\gamma \log(p_{y_i})`
+
+    Args:
+        gamma: :math:`\gamma > 0` puts more focus on hard misclassified samples
+
+    Shape:
+        - predictions :math:`(B, C)` : the logits
+        - target :math`(B, )` : the target ids to predict
+    """
+
+    def __init__(self, gamma):
+        super(FocalLoss, self).__init__()
+        self.gamma = gamma
+        self.eps = 1e-10
+
+    def forward(self, predictions, target):
+        probs = F.softmax(predictions, dim=1) + self.eps
+        target_one_hot = one_hot(
+            target,
+            num_classes=predictions.shape[1],
+            device=predictions.device,
+            dtype=predictions.dtype,
+        )
+
+        weight = torch.pow(1.0 - probs, self.gamma)
+        focal = -weight * torch.log(probs)
+        loss_tmp = torch.sum(target_one_hot * focal, dim=1)
+        loss = loss_tmp.mean()
+
+        return loss

### 2022-05-06 SampleCode/preprocess.py
doc string

@@ -36,6 +36,7 @@ import os
 
 # External modules
 import tqdm
+import torch
 
 # Local modules
 import data
@@ -43,9 +44,15 @@ import data
 IMAGE_IDX_FILEWIDTH = 6
 
 
-def process_dataset(dataset, outputdir):
+def process_dataset(dataset: torch.data.Dataset, outputdir: pathlib.Path):
+    """Loads and apply the transform on every image of a dataset
+
+    Args:
+        dataset (torch.data.Dataset) : a mappable dataset
+        outputdir (Pathlib.path) : an output dir to save the data in
+    """
     img_idx = {}
-    logging.info(f"Processsing {len(dataset)} objects in {outputdir}")
+    logging.info(f"Processing {len(dataset)} objects in {outputdir}")
     for X, y in tqdm.tqdm(dataset):
         if y not in img_idx:
             img_idx[y] = 0
@@ -68,6 +75,11 @@ def process_dataset(dataset, outputdir):
 
 
 def main(args):
+    """Precompte the training and validation images (resize them all)
+
+    Args:
+        args (dict): the parameters of the transform, data dir and output dir
+    """
     logging.info("Loading the images")
     train_dataset, valid_dataset = data.load_dataset(
         args.datadir,

### 2022-05-06 SampleCode/models.py
little code clean up

@@ -19,7 +19,6 @@
 
 import torch
 import torch.nn as nn
-import torchvision.models
 import torchvision.models.resnet as resnet
 import timm
 import timm.models.layers
@@ -83,8 +82,8 @@ class SimpleCNNRF(nn.Module):
 
     def forward(self, x):
         self.feature_maps = []
-        for l in self.features:
-            x = l(x)
+        for layer in self.features:
+            x = layer(x)
             self.feature_maps.append(x)
         x = x.view(x.size()[0], -1)
         return self.classifier(x)

### 2022-05-06 SampleCode/main.py
remove useless code, docstrings, factor mean/std, use ToTensorV2

@@ -36,7 +36,6 @@ import deepcs.display
 import deepcs.fileutils
 import torch
 import torch.nn as nn
-import torch.nn.functional as F
 import torch.optim
 from torch.utils.tensorboard import SummaryWriter
 import torchvision.transforms as transforms
@@ -51,8 +50,17 @@ import data
 import models
 import utils
 
+# Image mean and std can be computed by calling the "stats" function
+IMAGE_MEAN = 0.92
+IMAGE_STD = 0.16
+
 
 def stats(args):
+    """Computes the mean/std over all the pixels of the training data
+    The statistics are displayed in the console
+    Args:
+        args (dict): the parameters for loading the data
+    """
     print("Stats")
 
     use_cuda = torch.cuda.is_available()
@@ -86,38 +94,38 @@ def stats(args):
 
 
 def train(args):
-    print("Training")
+    """Train a neural network on the plankton classification task
+
+    Args:
+        args (dict): parameters for the training
+
+    Examples::
+
+        python3 main.py train --normalize --model efficientnet_b3
+    """
+    logging.info("Training")
 
     use_cuda = torch.cuda.is_available()
     device = torch.device("cuda") if use_cuda else torch.device("cpu")
 
-    # Get the data
-    # train_transforms = [
-    #     ImageOps.grayscale,
-    #     transforms.ToTensor()
-    # ]
-    # if args.normalize:
-    #     train_transforms.append(transforms.Normalize((0.92, ), (0.16, )))
-    # train_transforms.append(transforms.RandomRotation(180, fill=1.0))
-    # train_transforms = transforms.Compose(train_transforms)
-
+    # Set up the train and valid transforms
     train_transforms = [
+        ImageOps.GrayScale,
         A.HorizontalFlip(),
         A.VerticalFlip(),
         A.MotionBlur(),
-        A.CoarseDropout(fill_value=(255, 255, 255), max_height=20, max_width=20),
-        A.Rotate(180, p=0.5, border_mode=cv2.BORDER_CONSTANT, value=(255, 255, 255)),
+        A.CoarseDropout(fill_value=255, max_height=20, max_width=20),
+        A.Rotate(180, p=0.5, border_mode=cv2.BORDER_CONSTANT, value=255),
         data.ScaleBrightness(scale_range=(0.8, 1.0)),
-        data.KeepChannel(0, always_apply=True),
     ]
     if args.normalize:
[…]

### 2022-05-06 SampleCode/main.py
print <-> logging

@@ -61,7 +61,7 @@ def stats(args):
     Args:
         args (dict): the parameters for loading the data
     """
-    print("Stats")
+    logging.info("Computing the statistics over the training images")
 
     use_cuda = torch.cuda.is_available()
     device = torch.device("cuda") if use_cuda else torch.device("cpu")
@@ -90,7 +90,7 @@ def stats(args):
         )  # Mean over num_pixels
     std_pix /= num_imgs
     std_pix = math.sqrt(std_pix)
-    print(f"Mean pixel value : {mean_pix} (std {std_pix})")
+    logging.info(f"Mean pixel value : {mean_pix} (std {std_pix})")
 
 
 def train(args):
@@ -202,7 +202,7 @@ def train(args):
 
     with open(logdir / "summary.txt", "w") as f:
         f.write(summary_text)
-    print(summary_text)
+    logging.info(summary_text)
 
     # Callbacks
     tensorboard_writer = SummaryWriter(log_dir=logdir)
@@ -230,7 +230,7 @@ def train(args):
         test_metrics = deepcs.testing.test(model, valid_loader, device, val_metrics)
         macro_F1, class_F1 = utils.f1_metric(model, valid_loader, device)
         updated = model_checkpoint.update(macro_F1)
-        print(
+        logging.info(
             "[%d/%d] Test:   Loss : %.3f | F1 : %.3f | Acc : %.3f%% %s"
             % (
                 e,
@@ -242,7 +242,7 @@ def train(args):
             )
         )
 
-        print(f"Class F1 : {class_F1}")
+        logging.info(f"Class F1 : {class_F1}")
         # Confusion matrix
         cm = utils.make_confusion_matrix(model, valid_loader, device, num_classes)
         fig = utils.plot_confusion_matrix(cm)
@@ -349,7 +349,7 @@ if __name__ == "__main__":
     This is free software, and you are welcome to redistribute it
     under certain conditions;
     """
-    print(license)
+    logging.info(license)
 
     logging.basicConfig(level=logging.INFO, format="%(message)s")
     parser = argparse.ArgumentParser()

### 2022-05-08 SampleCode/data.py
allow stratified split

@@ -24,6 +24,7 @@ import os
 import argparse
 import tqdm
 import random
+from collections import defaultdict
 
 # External imports
 import torchvision
@@ -193,19 +194,39 @@ def load_raw_data(datadir):
     return MyDataset(root=datadir)
 
 
-def load_dataset(datadir, transform, val_ratio):
+def load_dataset(
+    datadir: pathlib.Path, transform: Callable, val_ratio: float, stratified: bool
+):
     dataset = MyDataset(root=datadir, transform=transform)
     logging.info(dataset)
 
-    indices = list(range(len(dataset)))
-    num_data = len(dataset)
-    num_valid = int(val_ratio * num_data)
-    num_train = num_data - num_valid
+    if not stratified:
+        indices = list(range(len(dataset)))
+        num_data = len(dataset)
+        num_valid = int(val_ratio * num_data)
+        num_train = num_data - num_valid
 
-    # TODO: split in valid/train with the same proportions
-    np.random.shuffle(indices)
-    train_indices = indices[:num_train]
-    valid_indices = indices[num_train:]
+        np.random.shuffle(indices)
+        train_indices = indices[:num_train]
+        valid_indices = indices[num_train:]
+    else:
+        # Browse all the dataset for giving the right proportion of samples
+        # per class
+        # As we do not know the number of classes, we use a dictionnary
+        class_indices = defaultdict(list)
+        for i, (_, yi) in enumerate(dataset):
+            class_indices[yi].append(i)
+
+        # Random assign the indices for the train and valid folds
+        train_indices, valid_indices = [], []
+        for ci, idxi in class_indices.items():
+            num_data = len(idxi)
+            num_valid = int(val_ratio * num_data)
+            num_train = num_data - num_valid
+
+            np.random.shuffle(idxi)
+            train_indices.append(indices[:num_train])
+            valid_indices.append(indices[num_train:])
 
     train_dataset = torch.utils.data.Subset(dataset, train_indices)
     valid_dataset = torch.utils.data.Subset(dataset, valid_indices)

### 2022-07-08 SampleCode/data.py
upd

@@ -328,7 +328,7 @@ def load_preprocessed_data(
         sample_weights, len(sample_weights)
     )
 
-    logging.info(f"Using a sampler with class_weights : {class_weights}")
+    logging.info(f"Using a sampler with sample_weights : {sample_weights}")
 
     train_loader = torch.utils.data.DataLoader(
         train_dataset,
@@ -520,6 +520,7 @@ class ScaleData(ImageOnlyTransform):
 
     def apply(self, img, **params):
         res = img / 255.0
+        print(res.dtype)
         return res
 
 

### 2022-07-08 SampleCode/main.py

@@ -67,19 +67,27 @@ def stats(args):
     # Get the data
     loaders = data.load_preprocessed_data(
         args.datadir,
-        train_transform=A.Compose([A.ToGray(), ToTensorV2()]),
-        valid_transform=data.__default_preprocessed_transform,
+        train_transform=A.Compose([A.ToGray(), data.ScaleData(), ToTensorV2()]),
+        valid_transform=A.Compose([A.ToGray(), data.ScaleData(), ToTensorV2()]),
         # val_ratio=args.val_ratio,
         batch_size=args.batch_size,
         num_workers=args.num_workers,
         pin_memory=use_cuda,
     )
     train_loader, valid_loader, n_samples_per_class = loaders
+
+    X, _ = next(iter(train_loader))
+    print(X)
+    print(X.dtype)
+
     mean_pix, std_pix, num_imgs = 0, 0, 0
+    print("go")
     for X, _ in train_loader:
+        print("go 1 ")
         X = X.to(device)
         mean_pix += X.shape[0] * X.mean().item()  # Mean over num_pixels
         num_imgs += X.shape[0]
+        print("go 2 ")
     mean_pix /= num_imgs  # Mean over num_pixels x batch_size
     for X, _ in train_loader:
         X = X.to(device)
@@ -111,8 +119,6 @@ def train(args):
         data.KeepChannel(0, always_apply=True),
         A.HorizontalFlip(),
         A.VerticalFlip(),
-        A.MotionBlur(),
-        A.CoarseDropout(fill_value=255, max_height=20, max_width=20),
         A.Rotate(180, p=0.5, border_mode=cv2.BORDER_CONSTANT, value=255),
         data.ScaleBrightness(scale_range=(0.8, 1.0)),
         data.ScaleData(always_apply=True),

### 2022-07-09 SampleCode/data.py
rm prints, fix stats dataloading

@@ -282,13 +282,8 @@ def load_preprocessed_data(
     train_dir = datadir / "train"
     valid_dir = datadir / "valid"
 
-    if isinstance(train_transform, A.core.composition.Compose):
-        print("Got albumentations augmented dataset")
-        train_dataset = torchvision.datasets.ImageFolder(train_dir)
-        train_dataset = AugmentedDataset(train_dataset, train_transform)
-    else:
-        print("Not albumentations")
-        train_dataset = torchvision.datasets.ImageFolder(train_dir, train_transform)
+    train_dataset = torchvision.datasets.ImageFolder(train_dir)
+    train_dataset = AugmentedDataset(train_dataset, train_transform)
 
     if label_smooth is not None:
         train_dataset = LabelSmoothedDataset(
@@ -475,7 +470,8 @@ class AugmentedDataset(torch.utils.data.Dataset):
     def __getitem__(self, idx):
         img, label = self.dataset[idx]
         img = np.array(img)
-        return self.transform(image=img)["image"], label
+        img = self.transform(image=img)["image"]
+        return img, label
 
     @property
     def samples(self):
@@ -520,7 +516,6 @@ class ScaleData(ImageOnlyTransform):
 
     def apply(self, img, **params):
         res = img / 255.0
-        print(res.dtype)
         return res
 
 

### 2022-07-09 SampleCode/main.py

@@ -67,8 +67,20 @@ def stats(args):
     # Get the data
     loaders = data.load_preprocessed_data(
         args.datadir,
-        train_transform=A.Compose([A.ToGray(), data.ScaleData(), ToTensorV2()]),
-        valid_transform=A.Compose([A.ToGray(), data.ScaleData(), ToTensorV2()]),
+        train_transform=A.Compose(
+            [
+                data.KeepChannel(0, always_apply=True),
+                data.ScaleData(always_apply=True),
+                ToTensorV2(),
+            ]
+        ),
+        valid_transform=A.Compose(
+            [
+                data.KeepChannel(0, always_apply=True),
+                data.ScaleData(always_apply=True),
+                ToTensorV2(),
+            ]
+        ),
         # val_ratio=args.val_ratio,
         batch_size=args.batch_size,
         num_workers=args.num_workers,
@@ -76,20 +88,13 @@ def stats(args):
     )
     train_loader, valid_loader, n_samples_per_class = loaders
 
-    X, _ = next(iter(train_loader))
-    print(X)
-    print(X.dtype)
-
     mean_pix, std_pix, num_imgs = 0, 0, 0
-    print("go")
-    for X, _ in train_loader:
-        print("go 1 ")
+    for X, _ in tqdm.tqdm(train_loader):
         X = X.to(device)
         mean_pix += X.shape[0] * X.mean().item()  # Mean over num_pixels
         num_imgs += X.shape[0]
-        print("go 2 ")
     mean_pix /= num_imgs  # Mean over num_pixels x batch_size
-    for X, _ in train_loader:
+    for X, _ in tqdm.tqdm(train_loader):
         X = X.to(device)
         std_pix += (
             X.shape[0] * ((X - mean_pix) ** 2).mean().item()

### 2022-07-10 Report/paper.tex
remove the neurips notice string
fix fig position
more on teams report

@@
 				Rotate($0, 360$), HFlip($0.5$), Translate($0.2$), Normalization$(0.5, 0.5$)&
-			None&
+			None& Custom CNNs, resnet34, VGG16, VGG19, regnet\_y\_32gf
 			&
-				Cross entropy loss, Adam($1e-4$), Weight decay $1e-4$, ..., batch size $128$, train/valid split&
+				Cross entropy loss, Adam($1e-4$), Weight decay $1e-4$, ..., batch size $128$, train/valid split (0.95, 0.05)&
 			batch sampler $f_k=1/count_k$\\
@@
 			The Backpropagated StuBBorns & 
-			Resize($224$), Normalize($0.5$, $0.5$) & 
+				SquarePad(), Resize($224$), Normalize($0.485$, $0.229$) & torchvision AutoAugment &None
+			& pretrained timm efficientnetv2\_rw\_s, tf\_efficientnet\_b8\_ap, deit\_tiny, cait\_s24\_224, coat\_mini, swin transformers, tnt\_s\_patch16\_224, dm\_nfnet, resnet50 followed by 2 linear/dropout/relu layers
+				& Cross Entropy Loss, Adam($1e-4$), early stopping on the validation loss, batch size $32$, train/valid split ($0.8, 0.2$) 
 			&
-			&
-			&
-			&
-			\\
+			batch sampler $f_k=1/count_k$\\
 			\hline
 			JMBmc & 
-			Resize($224$), Normalize($0.5$, $0.5$) & 
-			&
-			&
-			&
-			&
+				Resize($224$), 3 chans ImageNet Normalize & RandomRotate($30$), HFlip($0.5$), VFlip($0.5$)
+			& None
+				& pretrained timm tf\_efficient\_b4, pretrained torchvision, possibly frozen, feature extraction with resnet50, wideResnet50, mobilenet\_v2 followed by $5$ linear/dropout layers\footnote{unfortunately, in their implementation, the participants forgot to add non linearities between the layers, the classification stage is therefore only linear although composed of a stack of layers}
+				& Cross entropy loss, SGD($0.03$) with momentum, StepLR scheduler, batch size $64$, train/valid split ($0.8, 0.2$), early stopping on the validation loss
+			&None
 			\\
@@
 			AG0D & 
-			Resize($224$), Normalize($0.5$, $0.5$) & 
-			&
-			&
-			&
-			&
+				Resize($224$), 3 chans Imagenet Normalize & RandomRotate, RandomAffine, HFlip($0.5)$, VFlip($0.5$), 
+			& RandomRotate, RandomAffine, HFlip($0.5)$, VFlip($0.5$), 
+				& Custom CNNs, pretrained torchvision Resnet18, Resnet50, EfficientNetB7 ...
+with a random classification head\footnote{unfortunately, as JMBmc, the classifier is a stack of linear layers without non linearities}
+				& weighted cross entropy loss, Adam($0.001$), Reduce LR on plateau, batch size $64$, early stopping on the validation F1, train/valid split ($0.88, 0.12$), per class
+				& class weights in the CE loss $w_k=\log(\sum_i count_i/count_k) > 0$
 			\\
@@
 			ResMetz & 
-			Resize($224$), Normalize($0.5$, $0.5$) & 
+			SquarePad, Resize($224$), 3 chans ImageNet Normalize & 
+				HFlip($0.3$), VFlip($0.3$), RandomRotate
+			& None
+			& torchvision pretrained Resnet18, Resnet50, EfficientNet with a random classification head
+				& weighted cross entropy loss, Adam($3e-4$), early stopping the validation F1, train/valid split ($0.8, 0.2$)
 			&
-			&
-			&
-			&
+			class weights in the CE loss $w_k=\sqrt(\sum_i count_i/count_k)$, batch sampler with the same weight as the class weights
+
 			\\
@@
 
-\begin{figure}
+\begin{figure}[htbp]
 	\begin{center}

### 2022-07-10 SampleCode/main.py

@@ -122,11 +122,11 @@ def train(args):
     # Set up the train and valid transforms
     train_transforms = [
         data.KeepChannel(0, always_apply=True),
+        data.ScaleData(always_apply=True),
         A.HorizontalFlip(),
         A.VerticalFlip(),
-        A.Rotate(180, p=0.5, border_mode=cv2.BORDER_CONSTANT, value=255),
+        A.Rotate(180, p=0.5, border_mode=cv2.BORDER_CONSTANT, value=1.0),
         data.ScaleBrightness(scale_range=(0.8, 1.0)),
-        data.ScaleData(always_apply=True),
     ]
     if args.normalize:
         train_transforms.append(

### 2022-07-10 SampleCode/job.py
run some resnet18 and regnety_016 trainings without the CoarseDropout

@@ -118,11 +118,11 @@ commit_id = subprocess.check_output(
 os.system("mkdir -p logslurms")
 
 # Launch the batch jobs
-for model in ["resnet18"]:
+for model in ["resnet18", "regnety_016"]:
     submit_job(
         makejob(
             commit_id,
-            1,
+            2,
             "gpu_prod_long",
             "48:00:00",
             normalize=True,
@@ -135,7 +135,7 @@ for model in ["resnet18"]:
                 "nepochs": 100,
                 "base_lr": 0.001,
                 "loss": "BCE",
-                # "mixup": 0.2,
+                "mixup": 0.2,
             },
         )
     )

### 2022-08-08 SampleCode/preprocess.py
fix filepath to str as input to cv2 imwrite

@@ -73,7 +73,7 @@ def process_dataset(dataset: torch.utils.data.Dataset, outputdir: pathlib.Path):
         assert IMAGE_IDX_FILEWIDTH == 7
         output_filepath = output_filedir / f"{img_idx[y]:06d}.jpg"
         img_idx[y] += 1
-        cv2.imwrite(output_filepath, X)
+        cv2.imwrite(str(output_filepath), X)
 
 
 def main(args):

### 2022-08-09 SampleCode/job.py
fix cublas reproducibility variable

@@ -88,6 +88,9 @@ echo ""
 echo "Training"
 date
 
+# To enforce deterministic algorithms, cuda requires the following variable to be defined
+# see : https://docs.nvidia.com/cuda/cublas/index.html#cublasApi_reproducibility
+export CUBLAS_WORKSPACE_CONFIG=:4096:8
 python3 main.py  --datadir ./data/ {paramsstr} --logname {params['model']}_${{SLURM_ARRAY_JOB_ID}}_${{SLURM_ARRAY_TASK_ID}} --commit_id '{commit_id}' --logdir ${{current_dir}}/logs train
 
 if [[ $? != 0 ]]; then

### 2022-08-09 SampleCode/main.py
fix license program name

@@ -376,7 +376,7 @@ def test(args):
 if __name__ == "__main__":
     logging.basicConfig(stream=sys.stdout, level=logging.INFO, format="%(message)s")
     license = """
-    prepare_data.py  Copyright (C) 2022  Jeremy Fix
+    main.py  Copyright (C) 2022  Jeremy Fix
     This program comes with ABSOLUTELY NO WARRANTY;
     This is free software, and you are welcome to redistribute it
     under certain conditions;

### 2022-08-09 SampleCode/main.py
allow command line custom seed

@@ -383,11 +383,9 @@ if __name__ == "__main__":
     """
     logging.info(license)
 
-    utils.seed_everything()
-
-    parser = argparse.ArgumentParser()
     parser.add_argument("command", choices=["train", "test", "stats"])
 
+    parser.add_argument("--seed", type=int, default=42)
     parser.add_argument("--logdir", type=pathlib.Path, default="./logs")
     parser.add_argument("--commit_id", type=str, default=None)
     parser.add_argument("--datadir", type=pathlib.Path, required=True)
@@ -414,4 +412,9 @@ if __name__ == "__main__":
     parser.add_argument("--modelpath", type=pathlib.Path, default=None)
 
     args = parser.parse_args()
+
+    utils.seed_everything(args.seed)
+
+    parser = argparse.ArgumentParser()
+
     exec(f"{args.command}(args)")

### 2022-08-09 SampleCode/utils.py

@@ -215,7 +215,7 @@ class FocalLoss(nn.Module):
         return loss
 
 
-def seed_everything(seed=42):
+def seed_everything(seed):
     """Set the seed on pipeline, so the results are the same at every time we run it
     This function is for the reproducibility
     """

### 2022-08-09 SampleCode/job.py
fix parser bug, print seed, move nruns to 1

@@ -126,7 +126,7 @@ for model in ["regnety_016"]:
     submit_job(
         makejob(
             commit_id,
-            2,
+            1,
             "gpu_prod_long",
             "48:00:00",
             normalize=True,

### 2022-08-09 SampleCode/main.py

@@ -383,6 +383,8 @@ if __name__ == "__main__":
     """
     logging.info(license)
 
+    parser = argparse.ArgumentParser()
+
     parser.add_argument("command", choices=["train", "test", "stats"])
 
     parser.add_argument("--seed", type=int, default=42)
@@ -413,8 +415,7 @@ if __name__ == "__main__":
 
     args = parser.parse_args()
 
+    logging.info(f"Seeding the code with the seed {args.seed}")
     utils.seed_everything(args.seed)
-
-    parser = argparse.ArgumentParser()
 
     exec(f"{args.command}(args)")

### 2022-08-09 SampleCode/job.py
try multiple models with a fixed seed

@@ -122,7 +122,7 @@ commit_id = subprocess.check_output(
 os.system("mkdir -p logslurms")
 
 # Launch the batch jobs
-for model in ["regnety_016"]:
+for model in ["resnet18", "resnet152", "efficientnet_b3", "regnety_016"]:
     submit_job(
         makejob(
             commit_id,

### 2022-08-10 SampleCode/job.py
train efficientb3 and resnet152 with bs of 64

@@ -122,7 +122,7 @@ commit_id = subprocess.check_output(
 os.system("mkdir -p logslurms")
 
 # Launch the batch jobs
-for model in ["resnet18", "resnet152", "efficientnet_b3", "regnety_016"]:
+for model in ["resnet152", "efficientnet_b3"]:
     submit_job(
         makejob(
             commit_id,
@@ -134,7 +134,7 @@ for model in ["resnet18", "resnet152", "efficientnet_b3", "regnety_016"]:
             batch_sampler=True,
             params={
                 "model": model,
-                "batch_size": 128,
+                "batch_size": 64,
                 "weight_decay": 0.00,
                 "nepochs": 100,
                 "base_lr": 0.0003,

### 2022-08-11 SampleCode/main.py
fix test_transforms syntax

@@ -305,13 +305,15 @@ def test(args):
 
     test_transforms = data.__default_transform
     if args.normalize:
-        test_transforms.append(
-            A.Normalize(
-                (IMAGE_MEAN,), (IMAGE_STD,), max_pixel_value=1.0, always_apply=True
-            )
+        test_transforms = A.Compose(
+            [
+                test_transforms,
+                A.Normalize(
+                    (IMAGE_MEAN,), (IMAGE_STD,), max_pixel_value=1.0, always_apply=True
+                ),
+            ]
         )
-    test_transforms.append(ToTensorV2())
-    test_transforms = A.Compose(test_transforms)
+    test_transforms = A.Compose([test_transforms, ToTensorV2()])
 
     loader = data.load_test_data(
         args.datadir,

### 2022-08-11 SampleCode/job-infer.py
fix CUBLAS for deterministic

@@ -77,7 +77,8 @@ echo ""
 echo "Testing"
 date
 
-python3 main.py  --datadir {datadir} {paramsstr} test
+export CUBLAS_WORKSPACE_CONFIG=:4096:8
+python main.py  --datadir {datadir} {paramsstr} test
 
 if [[ $? != 0 ]]; then
     exit -1

### 2022-08-11 Report/paper.tex
description of the sample submission

@@
 
+\section{Sample submission}
+
+We provide a sample submission on the repository \url{https://github.com/jeremyfix/planktonChallenge}.
+This sample submission provides single models with reasonnable performances.
+Model averaging is not performed at the time of writting.
+
+\paragraph{Preprocessing}: Every image is square padded (with white color) and resized to $300 \times 300$, values are divided by $255$ and then normalized with mean $\mu=0.92$ and variance $0.16$.
+
+The training data are split in a training fold and a validation fold of $0.9/0.1$.
+The split is stratified meaning every class has $90\%$ of its samples in the training set and the remaining $10\%$ in the validation set.
+The sample code preprocesses and splits the data once for all making later simulations faster.
+
+\paragraph{Train augmentation}: the samples can be horizontally or vertically flipped, rotated (missing pixels are filled with white).
+CoarseDropout is also used to mask randomly parts of the image.
+Finally, as the images appeared to be sometimes blury, motion blur is added.
+
+%TODO : Add augmented samples
+
+There is no test time augmentation.
+
+\paragraph{Model architectures}: several models can be trained: from basic and custom linear and convolutional architectures to architectures provided by the Timm package (resnet18, resnet50, resnet152, efficientnet\_b3, efficientnet\_b3a, regnety\_016).
+All the models from Timm are initialized with weights pretrained on ImageNet.
+The first convolution and the last fully connected layers are randomly initialized to match the dimensions of the problem (one input channel, 86 output classes).
+
+\paragraph{Optimizer and scheduler}: The training loss is a cross entropy loss.
+The optimizer is Adam, with an initial learning rate of $0.0003$, a minibatch size of $128$ or $64$ (depending on the model size).
+The models are trained for $100$ epochs (although actually training time was limited to $48$ hours and large models do not perform more than $43$ epochs and smaller models such as resnet18 perform $100$ epochs in $35$ hours).
+
+Mixup regularization was also introduced.
+In mixup, every sample $(x, y)$ is built from two samples $(x_1, y_1)$ and $(x_2,y_2)$ where $x = \lambda x_1 + (1-\lambda)x_2$ and $y = \lambda y_1 + (1-\lambda y_2$ with $y_1, y_2$ the one-hot encoding of the classes and $\lambda$ sampled from a $\beta(0.2, 0.2)$ distribution.
+
+\paragraph{Class imbalance}: every minibatch is built by sampling the samples of the training folds with a weight equals to the inverse of the number of samples of the same class.
+
 %-----------------------------------------------------------------------------------

### 2022-08-13 SampleCode/data.py
fix : scalebrightness must take as input a max pixel vluae

@@ -462,9 +462,10 @@ class AugmentedDataset(torch.utils.data.Dataset):
 
 
 class ScaleBrightness(ImageOnlyTransform):
-    def __init__(self, scale_range, *args, **kwargs):
+    def __init__(self, scale_range, max_pixel_value, *args, **kwargs):
         super().__init__(*args, **kwargs)
         self.scale_range = scale_range
+        self.max_pixel_value = max_pixel_value
 
     def get_transform_init_args_names(self):
         return ["scale_range"]
@@ -473,7 +474,7 @@ class ScaleBrightness(ImageOnlyTransform):
         f = self.scale_range[0] + np.random.random() * (
             self.scale_range[1] - self.scale_range[0]
         )
-        return 255 - np.uint8((255 - img) * f)
+        return self.max_pixel_value - np.uint8((self.max_pixel_value - img) * f)
 
 
 class ScaleData(ImageOnlyTransform):

### 2022-08-13 SampleCode/main.py

@@ -127,7 +127,7 @@ def train(args):
         A.MotionBlur(),
         A.CoarseDropout(fill_value=1.0, max_height=20, max_width=20),
         A.Rotate(180, p=0.5, border_mode=cv2.BORDER_CONSTANT, value=1.0),
-        data.ScaleBrightness(scale_range=(0.8, 1.0)),
+        data.ScaleBrightness(scale_range=(0.8, 1.0), max_pixel_value=1.0),
     ]
     if args.normalize:
         train_transforms.append(

### 2022-08-13 SampleCode/data.py
fix test_augmentations, change default data tests datadir

@@ -503,37 +503,28 @@ class KeepChannel(ImageOnlyTransform):
 
 
 def test_augmentations(args):
-    num_imgs = 36
+    num_img = 36
     transform = A.Compose(
         [
+            __default_transform,
             A.HorizontalFlip(),
             A.VerticalFlip(),
             A.MotionBlur(),
-            A.CoarseDropout(fill_value=(255, 255, 255), max_height=20, max_width=20),
-            A.Rotate(
-                180, p=0.5, border_mode=cv2.BORDER_CONSTANT, value=(255, 255, 255)
-            ),
+            A.CoarseDropout(fill_value=1.0, max_height=20, max_width=20),
+            A.Rotate(180, p=0.5, border_mode=cv2.BORDER_CONSTANT, value=1.0),
             ScaleBrightness(scale_range=(0.8, 1.0)),
-            KeepChannel(0, always_apply=True),
-            A.Normalize((0.92,), (0.16,)),
-            ToTensorV2(),
+            # A.Normalize((0.92,), (0.16,)),
+            # ToTensorV2(),
         ]
     )
-    print(type(transform))
-    print(isinstance(transform, A.core.composition.Compose))
-    images = []
-    fnames = os.scandir(args.datadir)
-    for img in fnames:
-        imgpath = os.path.join(args.datadir, img.name)
-        img = cv2.imread(imgpath)
-        images.append(transform(image=img)["image"])
-        if len(images) == num_imgs:
-            break
+    dataset = MyDataset(root=args.datadir, transform=transform)
 
     fig = plt.figure()
-    for i, img in enumerate(images):
+    for i in range(6):
+        img, _ = dataset[num_img]
+        print(img.min(), img.max())
         ax = fig.add_subplot(6, 6, i + 1)
-        ax.imshow(img[0], cmap="gray")
+        ax.imshow(img, cmap="gray")
         ax.axis("off")
     plt.show()
 
@@ -650,7 +641,7 @@ if __name__ == "__main__":
     parser = argparse.ArgumentParser()
 
     parser.add_argument(
-        "--datadir", type=pathlib.Path, default="/opt/Datasets/ZooScanSet/imgs"
+        "--datadir", type=pathlib.Path, default="/mounts/Datasets1/ChallengeDeep/test"
     )
     parser.add_argument("--batch_size", default=64)
     parser.add_argument("--classidx", type=int, default=None)
@@ -659,11 +650,11 @@ if __name__ == "__main__":
 
[…]

### 2022-08-13 SampleCode/data.py
remove motion blur, fix SquareResize pad mode, fix scale brightness max value, update test_augmentations code

@@ -61,6 +61,7 @@ class SquareResize(ImageOnlyTransform):
             A.PadIfNeeded(
                 min_height=longest_size,
                 min_width=longest_size,
+                border_mode=cv2.BORDER_CONSTANT,
                 value=255,
                 always_apply=True,
             ),
@@ -474,7 +475,9 @@ class ScaleBrightness(ImageOnlyTransform):
         f = self.scale_range[0] + np.random.random() * (
             self.scale_range[1] - self.scale_range[0]
         )
-        return self.max_pixel_value - np.uint8((self.max_pixel_value - img) * f)
+        return (self.max_pixel_value - (self.max_pixel_value - img) * f).astype(
+            img.dtype
+        )
 
 
 class ScaleData(ImageOnlyTransform):
@@ -503,16 +506,15 @@ class KeepChannel(ImageOnlyTransform):
 
 
 def test_augmentations(args):
-    num_img = 36
     transform = A.Compose(
         [
             __default_transform,
             A.HorizontalFlip(),
             A.VerticalFlip(),
-            A.MotionBlur(),
+            # A.MotionBlur(),
             A.CoarseDropout(fill_value=1.0, max_height=20, max_width=20),
             A.Rotate(180, p=0.5, border_mode=cv2.BORDER_CONSTANT, value=1.0),
-            ScaleBrightness(scale_range=(0.8, 1.0)),
+            ScaleBrightness(scale_range=(0.8, 1.0), max_pixel_value=1.0),
             # A.Normalize((0.92,), (0.16,)),
             # ToTensorV2(),
         ]
@@ -521,12 +523,17 @@ def test_augmentations(args):
 
     fig = plt.figure()
     for i in range(6):
-        img, _ = dataset[num_img]
-        print(img.min(), img.max())
-        ax = fig.add_subplot(6, 6, i + 1)
-        ax.imshow(img, cmap="gray")
+        img, _ = dataset.imgdataset.dataset[i]
+        ax = fig.add_subplot(6, 6, 6 * i + 1)
+        ax.imshow(img, cmap="gray", clim=[0, 1])
         ax.axis("off")
-    plt.show()
+        for j in range(5):
+            img, _ = dataset[i]
+            ax = fig.add_subplot(6, 6, 6 * i + j + 2)
+            ax.imshow(img, cmap="gray", clim=[0, 1])
+            ax.axis("off")
+    plt.savefig("augmented_images.pdf", bbox_inches="tight")
+    # plt.show()
 
 
 def test_augmented_dataloader(args):

### 2022-08-13 SampleCode/job.py
runs with the fixed augmentated data

@@ -122,7 +122,7 @@ commit_id = subprocess.check_output(
 os.system("mkdir -p logslurms")
 
 # Launch the batch jobs
-for model in ["resnet152", "efficientnet_b3"]:
+for model in ["resnet18", "regnety_016", "resnet152", "efficientnet_b3"]:
     submit_job(
         makejob(
             commit_id,

### 2022-08-13 SampleCode/main.py
remove motion blur in the training augmentations

@@ -124,7 +124,7 @@ def train(args):
         data.__default_transform,
         A.HorizontalFlip(),
         A.VerticalFlip(),
-        A.MotionBlur(),
+        # A.MotionBlur(),
         A.CoarseDropout(fill_value=1.0, max_height=20, max_width=20),
         A.Rotate(180, p=0.5, border_mode=cv2.BORDER_CONSTANT, value=1.0),
         data.ScaleBrightness(scale_range=(0.8, 1.0), max_pixel_value=1.0),

### 2022-08-13 SampleCode/job.py
fix default seed to None in the codes, to 42 in the slurm scripts

@@ -140,6 +140,7 @@ for model in ["resnet18", "regnety_016", "resnet152", "efficientnet_b3"]:
                 "base_lr": 0.0003,
                 "loss": "BCE",
                 "mixup": 0.2,
+                "seed": 42,
             },
         )
     )

### 2022-08-13 SampleCode/main.py

@@ -389,7 +389,7 @@ if __name__ == "__main__":
 
     parser.add_argument("command", choices=["train", "test", "stats"])
 
-    parser.add_argument("--seed", type=int, default=42)
+    parser.add_argument("--seed", type=int, default=None)
     parser.add_argument("--logdir", type=pathlib.Path, default="./logs")
     parser.add_argument("--commit_id", type=str, default=None)
     parser.add_argument("--datadir", type=pathlib.Path, required=True)
@@ -417,7 +417,8 @@ if __name__ == "__main__":
 
     args = parser.parse_args()
 
-    logging.info(f"Seeding the code with the seed {args.seed}")
-    utils.seed_everything(args.seed)
+    if args.seed is not None:
+        logging.info(f"Seeding the code with the seed {args.seed}")
+        utils.seed_everything(args.seed)
 
     exec(f"{args.command}(args)")

### 2022-08-13 SampleCode/preprocess.py

@@ -98,16 +98,18 @@ def main(args):
 if __name__ == "__main__":
     logging.basicConfig(level=logging.INFO, format="%(message)s")
 
-    utils.seed_everything()
-
     parser = argparse.ArgumentParser()
     parser.add_argument("--datadir", type=pathlib.Path, required=True)
     parser.add_argument("--outputdir", type=pathlib.Path, required=True)
     parser.add_argument("--num_workers", type=int, default=4)
     parser.add_argument("--batch_size", type=int, default=64)
     parser.add_argument("--val_ratio", type=float, default=0.05)
+    parser.add_argument("--seed", type=int, default=None)
 
     args = parser.parse_args()
+
+    if args.seed is not None:
+        utils.seed_everything(args.seed)
     if args.outputdir.exists():
         logging.info("The datadir exists, nothing to be done")
     else:

### 2022-08-13 SampleCode/data.py
fix PadIfNeeded : a single value is considered as the blue channel and not applied to all the chanels

@@ -56,13 +56,19 @@ class SquareResize(ImageOnlyTransform):
 
     def apply(self, img, **params):
         longest_size = max(*img.shape[:2])  # img is (H, W, C)
-        img = img.astype(np.float32)
+
+        # Super important
+        # PadIfNeeded value must be 3 channels
+        # torchvsiion ImageFolder loads grayscale images as 3 channels
+        # PadIfNeeded calls internally copyMakeBorder from cv2
+        # If you give value = 255 to copyMakeBorder, it will pad in blue
+        # not with the same value for the 3 channels which I find weird.
         transforms = [
             A.PadIfNeeded(
                 min_height=longest_size,
                 min_width=longest_size,
                 border_mode=cv2.BORDER_CONSTANT,
-                value=255,
+                value=(255, 255, 255),
                 always_apply=True,
             ),
             A.Resize(
@@ -70,7 +76,8 @@ class SquareResize(ImageOnlyTransform):
             ),
         ]
         transform = A.Compose(transforms)
-        return transform(image=img)["image"]
+        transformed_image = transform(image=img)["image"]
+        return transformed_image
 
 
 class MyDataset(torch.utils.data.Dataset):

### 2022-08-16 Report/scripts/plot_best_submissions.py
best submission output is now the executed dir ; the generated figures/tex must be copied manually

@@ -93,7 +93,7 @@ if __name__ == "__main__":
     plt.xlabel("Team name")
     plt.ylabel("Class F1 scores")
     plt.xticks(rotation=45, ha="right")
-    plt.savefig("../figs/class_f1_scores_per_team.pdf", bbox_inches="tight")
+    plt.savefig("class_f1_scores_per_team.pdf", bbox_inches="tight")
 
     print(class_stats[teams].T.head())
 
@@ -102,7 +102,7 @@ if __name__ == "__main__":
     plt.xlabel("Class")
     plt.ylabel("Class F1 scores")
     plt.xticks(rotation=45, ha="center")
-    plt.savefig("../figs/class_f1_scores_per_class.pdf", bbox_inches="tight")
+    plt.savefig("class_f1_scores_per_class.pdf", bbox_inches="tight")
 
     # Get the best predicted classes and the worst ones
     stats = class_stats[teams].T.describe().T
@@ -116,7 +116,7 @@ if __name__ == "__main__":
     print(f"The min F1 over all the teams being :\n {best_predicted_f1}")
 
     # Generate and save the latex tables to be included in the paper
-    with open("../figs/table_best_predicted.tex", "w") as f:
+    with open("table_best_predicted.tex", "w") as f:
         f.write(
             """
 \\begin{table}
@@ -139,7 +139,7 @@ $0.8$.}
 \\end{table}
                 """
         )
-    with open("../figs/table_worst_predicted.tex", "w") as f:
+    with open("table_worst_predicted.tex", "w") as f:
         f.write(
             """
 \\begin{table}

### 2022-08-16 Report/paper.tex
upd
images for the samples submission, scores tables, ..

@@
 \paragraph{Train augmentation}: the samples can be horizontally or vertically flipped, rotated (missing pixels are filled with white).
-CoarseDropout is also used to mask randomly parts of the image.
-Finally, as the images appeared to be sometimes blury, motion blur is added.
+Coarse dropout, where little square regions are filled in white, is also used to mask randomly parts of the image.
+Examples of augmented images are shown on figure~\ref{fig:sample_submission_augmented}.
 
-%TODO : Add augmented samples
+\begin{figure}[htbp]
+	\begin{center}
+	\includegraphics[width=0.75\columnwidth]{figs/augmented_images.pdf}
+	\caption{\label{fig:sample_submission_augmented} Examples of augmented images from the sample submission.
+The transforms are Horizontal flip, vertical flip, random rotation and coarse dropout.
+The first column represents the original images and the five following columns examples of augmented images.
+The little white squares on the images show the coarse dropout.}
+	\end{center}
+\end{figure}
 
@@
 \paragraph{Class imbalance}: every minibatch is built by sampling the samples of the training folds with a weight equals to the inverse of the number of samples of the same class.
+
+The cross entropy loss for the training set and the F1 score for the validation split are shown on figure~\ref{fig:sample_submission_losses}.
+The confusion matrices of the best parameters for the four trained networks are shown on figure~\ref{fig:sample_submission_confusion_matrix}.
+For every run, the best parameters are the one with the largest validation F1 score.
+
+\begin{figure}[!htbp]
+	\caption{\label{fig:sample_submission_losses} The cross entropy loss on the training set and the macro F1 score on the validation fold for the four trained models.}
+\end{figure}
+
+\begin{figure}[htbp]
+	\begin{center}
+		\raisebox{-0.5\height}{\includegraphics[width=0.25\columnwidth]{figs/class_f1_scores_per_team_sample.pdf}}
+		\hspace*{.4in}
+		\raisebox{-0.5\height}{\includegraphics[width=0.62\columnwidth]{figs/class_f1_scores_per_class_sample.pdf}}
+	\end{center}
+	\caption{\label{fig:class_team_score_sample} Class F1 scores on the test data per model (left) and per class (right).}
+\end{figure}
+
+\begin{figure}[!htbp]
+	\begin{tabular}{ccccc}
+		\includegraphics[width=0.18\columnwidth]{figs/confusion_matrix_resnet18.png}&
+		\includegraphics[width=0.18\columnwidth]{figs/confusion_matrix_resnet152.png}&
+		\includegraphics[width=0.18\columnwidth]{figs/confusion_matrix_efficientnetb3.png}&
+		\includegraphics[width=0.18\columnwidth]{figs/confusion_matrix_regnety16.png} &
+		\includegraphics[width=0.18\columnwidth]{figs/confusion_matrix_average.png}\\
+		Resnet18 & Resnet152 & EfficientNet B3 & Regnety 16 & Average model
+	\end{tabular}
+	\caption{\label{fig:sample_submission_confusion_matrix} Confusion matrices on the validation fold of the best parameters for the four trained models.}
+\end{figure}
+
+The macro F1 scores on the public and private test sets of the four trained models are given in table~\ref{table:sample_submission_scores}.
+We also provide the score for the entry which predicts the class as the one maximizing the average probability of the four models.
+The class F1 scores for each sample submission per model and per class are displayed on the figure~\ref{fig:class_team_score_sample}.
+
+\begin{table}[htbp]
+	\begin{center}
+	\begin{tabular}{|c|c|c|}
+		\hline
+		Model & Public macro F1 & Private macro F1 \\
+		\hline
+		Resnet18 & 0.79723 & 0.79344 \\
+		Resnet152 & 0.80178 & 0.79714 \\
+		EfficientNet b3 & 0.80041 & 0.80489\\
+		Regnety 16 & 0.80178 & 0.79714 \\
+		\hline\hline
+		Average model & 0.82687 & 0.82997\\
+		\hline
+	\end{tabular}
+	\end{center}
+	\caption{\label{table:sample_submission_scores} Macro F1 scores on the public and private test sets of the challenge.}
+
+\end{table}
 

### 2022-08-16 Report/paper.tex
rm useless figure

@@
 
-The cross entropy loss for the training set and the F1 score for the validation split are shown on figure~\ref{fig:sample_submission_losses}.
+The cross entropy loss for the training and validation sets and the F1 score for the validation split are shown on figure~\ref{fig:sample_submission_train_CE}-\ref{fig:sample_submission_test_F1}.
 The confusion matrices of the best parameters for the four trained networks are shown on figure~\ref{fig:sample_submission_confusion_matrix}.
@@
 
-\begin{figure}[!htbp]
-	\caption{\label{fig:sample_submission_losses} The cross entropy loss on the training set and the macro F1 score on the validation fold for the four trained models.}
+\begin{figure}
+	\centering
+	\def\svgwidth{0.5\columnwidth}
+	\input{figs/metrics_train_CE.pdf_tex}
+	\caption{\label{fig:sample_submission_train_CE}Averaged train cross entropy.Each were allowed to run for 48 hours.
+Light blue is resnet18, dark blue is resnet152, orange is regnety\_16 and red is efficientnet b3.
+}
+\end{figure}
+\begin{figure}
+	\centering
+	\def\svgwidth{0.5\columnwidth}
+	\input{figs/metrics_test_CE.pdf_tex}
+	\caption{\label{fig:sample_submission_valid_CE}Averaged cross entropy on the validation fold.Each were allowed to run for 48 hours.
+Light blue is resnet18, dark blue is resnet152, orange is regnety\_16 and red is efficientnet b3.
+}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\def\svgwidth{0.5\columnwidth}
+	\input{figs/metrics_fix_test_F1.pdf_tex}
+	\caption{\label{fig:sample_submission_test_F1}Macro F1 scores on the validation fold for the four trained sample submissions.
+Each were allowed to run for 48 hours.
+Light blue is resnet18, dark blue is resnet152, orange is regnety\_16 and red is efficientnet b3.
+}
 \end{figure}

### 2022-08-16 Report/paper.tex
rm useless comment
last team descriptions

@@
 			CelestineFeuillat & 
-			Resize($224$), Normalize($0.5$, $0.5$) & 
-			&
-			&
-			&
-			&
+				Resize($224$), RandAugment(num\_ops=4, magnitude=7, fill=255), ImageNet normalize 
+			& None 
+			& None
+			& LinearNet, custom CNN, torchvision pretrained resnet18, resnet50
+				& Weighted cross entropy() loss, Adam($0.001$),early stopping on validation F1, train/valid random split ($0.9, 0.1$), batch size $128$
+				&class weights in the CE loss $w_k=\frac{\sum_i count_i}{86 count_k})$,
 			\\
@@
 			LesJongleurs & 
-			Resize($224$), Normalize($0.5$, $0.5$) & 
-			&
-			&
-			&
-			&
+			Resize($224$), Normalize($0.5$, $0.5$) 
+			&  None
+			& None
+			& LinearNet, torchvision pretrained resnet18, resnet50 
+			& weighted cross entropy, Adam($0.001$), early stopping on validation F1, train/valid random split ($0.8, 0.2$)
+			&class weights in the CE loss $w_k=\frac{\sum_i count_i}{count_k})$,
 			\\
@@
 			DeepBeru4s & 
-			Resize($224$), Normalize($0.5$, $0.5$) & 
-			&
-			&
-			&
-			&
+			Resize($256$), CenterCrop($224$), Resize($256$), ImageNet Normalize 
+			& HFlip ($0.5$), RandomRotate ($-15,15$)
+			& HFlip ($0.5$), RandomRotate ($-15,15$)
+			&  Linear, custom CNN, torchvision pretrained resnet34, resnet50, resnet152 with a possibly frozen backbone
+				& weighted cross entropy loss or F1 loss\footnote{actually incorrectly coded}, Adam($0.001$), early stopping on validation weighted cross entropy, train/valid  random split ($0.8, 0.2$), batch size $128$
+			&class weights in the CE loss $w_k=\frac{86}{count_k})$,batch sampler with $f_k = 1/count_k$
 			\\
 			\hline
-			The Whale Sharks & 
-			Resize($224$), Normalize($0.5$, $0.5$) & 
-			&
-			&
-			&
-			&
+			The Whale Sharks 
+				& Invert, CenterCrop($300$) or Resize($300$), Normalize($0.013$,$0.0546$)
+				& HFlip($0.5$), RandomRotate, Blur,
+			& None
+			& custom CNN, torchvision pretrained resnet101, resnet152, vgg16, vgg19
+				& cross entropy loss or micro average F1 loss, Adam($0.001$), early stopping on the batch averaged micro average F1, train/valid random split ($0.8, 0.2$), batch size $256$ 
+				& batch sampler with $f_k = 1/count_k$
 			\\
@@
 For example, looking at the samples provided in appendix, the class 33 "Nectophore Abylopsitetragona" seems pretty similar to the class 52 "Nectophore Diphyidae".
-
-%TODO : check why the off diagonal elements class do not match the badly classified classes listed above. There seems to be issues for class 11 for example ..
 
